### 2015-11-28 paper.tex
ER finishes first run through of edits
some questions included for DLM included in italics in .rmd

\documentclass[article]{jss}
\usepackage[utf8]{inputenc}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\author{
David L Miller\\University of St Andrews \And Eric Rexstad\\University of St Andrews \And Jeffrey L Laake\\National Marine Mammal Laboratory
}
\title{Distance Sampling in \proglang{R}}
\Keywords{distance sampling, abundance estimation, line transects, point transects, \proglang{R}}

\Abstract{
The abstract of the article.
}

\Plainauthor{David L Miller, Eric Rexstad, Jeffrey L Laake}
\Plaintitle{Distance Sampling in R}
\Plainkeywords{distance sampling, abundance estimation, line transects, point transects, R}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
\Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
    David L Miller\\
  University of St Andrews\\
  Centre for Research into Ecological and Environmental Modelling, The
  Observatory, St Andrews, Fife KY16 9LZ, Scotland\\
  E-mail: \href{mailto:dave@ninepointeightone.net}{\nolinkurl{dave@ninepointeightone.net}}\\
  URL: \url{http://converged.yt}\\~\\
      Eric Rexstad\\
  University of St Andrews\\
  Centre for Research into Ecological and Environmental Modelling, The
  Observatory, St Andrews, Fife KY16 9LZ, Scotland\\
  E-mail: \href{mailto:Eric.Rexstad@st-andrews.ac.uk}{\nolinkurl{Eric.Rexstad@st-andrews.ac.uk}}\\

      Jeffrey L Laake\\
  National Marine Mammal Laboratory\\
  Alaska Fisheries Science Center 7600 Sand Point Way N.E., Seattle, WA
  98115, USA\\
  E-mail: \href{mailto:Jeff.Laake@noaa.gov}{\nolinkurl{Jeff.Laake@noaa.gov}}\\

  }

\usepackage{amsmath} \usepackage{amssymb} \usepackage{longtable}
\usepackage{booktabs}

\begin{document}

\section{Introduction}\label{introduction}

Distance sampling (Buckland et al. 2001; Buckland et al. 2004)
encompasses a suite of field methods and statistical models used to
estimate the abundance of biological populations. Distance sampling
field procedure can be thought of as an extension of plot sampling,
where we wish to take into account the decreasing probability of
detecting objects at increasing distance from the sampler. We do this by
building a model for detectability and use quantities calculated from
the detection function to adjust the observed counts to obtain an
estimate of abundance.

For many years distance sampling analyses have been available via the
Windows program Distance (or ``DISTANCE''); for clarity henceforth
``Distance for Windows'' Thomas et al. (2010)). From version 5 of
Distance for Windows, R packages have been included to perform
particular analyses (CITE Distance user manual). This paper shows how to
fit detection functions, perform model checking and selection and
estimate abundance in \proglang{R} using the package \pkg{Distance}.
\pkg{Distance} is a wrapper package around the more complex (and more
powerful) \pkg{mrds} and offers a subset of the analyses possible with
that package.

\subsection{Distance sampling}\label{distance-sampling}

Census-type surveys (quadrat or strip transects assuming perfect
detectability within strips or quadrats) are inefficient (requiring
considerable field effort) and we should expect that not all objects
(animals, plants, dung, etc) can be observed. Accounting for imperfect
detectability is an important consideration when obtaining accurate
estimates of abundance (Lahoz-Monfort, Guillera-Arroita, and Wintle
2013). Using the extra information gained by recording distance from the
sampler to the observation, it is possible to model detectability.
Because we expect detectability to decrease with increasing distance
from the sampler, we model detectability as a function of distance (plus
perhaps other covariates, see below). We use the probabilities of
detection calculated from the detection function to estimate an
abundance for the area covered by the survey, which we then scale-up to
the study region of interest.

Distance sampling surveys are conducted using two transect types: line
and point transects. In line transect sampling observers walk (or fly,
sail, etc) lines observing objects and recording the distances to the
line; whereas in point transect sampling observers remain stationary at
a location and record distances from that point. Field methods are
chosen to be suitable to species and habitat constraints (Buckland et
al. 2015).

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/points-and-lines-1}

}

\caption{Left side plots show an example of a survey of an area containing a population of 500 objects, blue indicates sampler placement (top lines, bottom points) and red dots indicate detected individuals. The right side of the figure shows histograms of observed distances (again, lines top and points bottom).\label{fig:pointslines}}\label{fig:points-and-lines}
\end{figure}
\end{CodeChunk}

For both points and lines, given the geometry of the sampler (see
``Detection functions''), the histogram of distances should show a
decreasing number of observations with increasing distance from the
sampler. For line transects we expect objects to be distributed
uniformly with respect to distance from the line with detectability
causing the number of detections to decrease. For point transects, we
recognise that as distance from the point increases, the area of the
circle encompassed increases with distance squared; hence the number of
objects available to be detected is a linearly increasing function of
distance from the point.

Using this histogram we can crudely estimate the drop-off in
detectability by eye by tracing a line that approximates the tops of the
histogram bars -- this is the detection function. \pkg{Distance}
estimates the parameters for a fixed-form detection function using
maximum likelihood estimation. We address possible models below.

Figure \ref{fig:pointslines} shows sampling of a population of 500
individuals using line and point transects (left column) and their
corresponding histograms (right column).

\subsection{Data}\label{data}

We demonstrate \pkg{Distance} using two data sets: one line transect and
one point transect. These data sets have been chosen to be
representative of data seen in practice.

\subsubsection{Minke whales}\label{minke-whales}

The line transect data is based on a survey of Antarctic minke whales
(\emph{Balaenoptera bonaerensis}). The data are simulated from models
fitted to data from the International Whaling Commission's International
Decade of Cetacean Research Southern Ocean Whale and Ecosystem Research
(IWC IDCR-SOWER) programme 1992-1993 austral summer surveys (Branch and
Butterworth 2001). They consist of 99 observations and include
geographical strata (near or distant from ice edge) and transect lengths
on the effort expended.

\subsubsection{Amakihi}\label{amakihi}

The point transect data set consists of 1485 observations of Amakihi
(\emph{Hemignathus virens}; a Hawaiian songbird), collected at 41 points
between 1992 and 1995. The data include distances and two covariates
collected during the survey: observer (a three level factor), time after
sunrise (transformed to minutes (continuous) or hours (factor)
covariates). Data are analysed comprehensively in Marques et al. (2007).

\subsection{The rest of the paper}\label{the-rest-of-the-paper}

The rest of the paper has this structure: we describe data organisation
for use with \pkg{Distance}; models for the detection function are
described in terms of formulation and examples of fitting in
\proglang{R}. We then examine model checking, goodness of fit and model
selection. Having illustrated how to obtain a good detection function,
we show how to estimate abundance using that model, including
stratification. The final two sections of the article look at extensions
(both in terms of methodology and software) and put the package in a
broader context amongst other software packages used for estimating the
abundance of biological populations.

\section{Data setup}\label{data-setup}

The two example data sets used here are distributed with \pkg{Distance}
so readers can reproduce our results. However, data will be collected in
the field and formatted correctly for use with \pkg{Distance}. The
package allows for a flexible format for data input ranging from very
simple to complex:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  In the simplest case, where the objective is to estimate a detection
  function, all that is needed is a vector of distances.
\item
  To include additional covariates into the detection function (see
  ``Detection functions'') we use a \code{data.frame}. The data.frame
  contains a column called \code{distance} and additional named columns
  for covariates potentially useful in modelling detectability (for
  example \code{observer} or \code{seastate}). Some column names are
  reserved: \code{object} for an observation identifier (see
  ``Extensions''), \code{size} for group or cluster size (see
  ``Detection functions'' and ``Abundance and variance estimation''),
  \code{detected} for whether an observation was detected (see
  ``Extensions'') and the columns described in the next bullet.
\item
  To estimate abundance beyond area covered by expended survey effort,
  additional information is required. This consists of information about
  which transect and occasion the observation was made (the
  \code{Sample.Label}), a column named \code{Effort} which gives the
  effort associated with that sample (for lines their length and for
  points the number of times that point was visited), the stratum the
  sample was located in (this may have any name and may be from pre- or
  post-survey stratification, see ``Estimating abundance and variance'')
  and strata areas (which has the same name as the stratum column,
  appended with \code{.Area}). We refer to this data format as
  ``flatfile'' where all information is contained in one table.
\end{itemize}

As we will see in ``Extensions'', further information is also required
when using more complex models.

The minke whale data follows the ``flatfile'' format given in the last
bullet point:

\begin{CodeChunk}
\begin{CodeInput}
library(Distance)
head(minke)
\end{CodeInput}
\begin{CodeOutput}
  Region.Label  Area Sample.Label Effort distance
1        South 84734            1  86.75     0.10
2        South 84734            1  86.75     0.22
3        South 84734            1  86.75     0.16
4        South 84734            1  86.75     0.78
5        South 84734            1  86.75     0.21
6        South 84734            1  86.75     0.95
\end{CodeOutput}
\end{CodeChunk}

Whereas the amakihi data lacks effort and stratum data:

\begin{CodeChunk}
\begin{CodeInput}
head(amakihi)
\end{CodeInput}
\begin{CodeOutput}
   survey object distance obs mas has detected
1 July 92      1       40 TJS  50   1        1
2 July 92      2       60 TJS  50   1        1
3 July 92      3       45 TJS  50   1        1
4 July 92      4      100 TJS  50   1        1
5 July 92      5      125 TJS  50   1        1
6 July 92      6      120 TJS  50   1        1
\end{CodeOutput}
\end{CodeChunk}

We will explore below the effects of including effort data during
analysis.

\section{Detection functions}\label{detection-functions}

The detection function describes the relationship between observed
distances and probability of detection. The detection function itself
models the probability
\(\mathbb{P}(\text{object detected } \vert \text{ object at distance } y)\)
and is usually denoted \(g(y; \boldsymbol{\theta})\) where \(y\) is
distance (either from a line or point) and \(\boldsymbol{\theta}\) is a
vector of parameters to be estimated. Our goal is to estimate an
\emph{average probability of detection} (\(p\), average in the sense of
an average over distances), so we must integrate out distance (\(y\))
from the detection function: \[
p = \int_0^w \pi(y) g(y; \boldsymbol{\theta}) dy
\] where \(\pi(y)\) describes the distribution of objects with respect
to the sampler and \(\pi(x)=1/w\) for line transects and
\(\pi(r)=\frac{2r}{w^2}\) for point transects, taking into account the
geometry of the sampler (Buckland et al. 2001, Chapter 3) (letting \(x\)
denote a perpendicular distance from a line and \(r\) denote radial
distance from a point).

It is crucial that the detection function accurately models
detectability at small distances. We are less worried by its behaviour
further away from 0. To ensure that the model is not overly influence by
distances far from zero, we truncate the distances beyond a given
distance \(w\) (known as the \emph{truncation distance}). Fitting to
distances at great distances from zero does not demonstrably improve the
precision of abundance estimates (Buckland et al. 2001, 103--8,
151--53).

Models for the detection function are expected to have the following
properties (Buckland et al. 2015, Chapter 5):

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \emph{Shoulder}: we expect observers to be able to see objects near
  them, not just those directly adjacent to them. For this reason, we
  expect the detection function to be horizontal near zero distance.
\item
  \emph{Non-increasing}: we do not think that observers should be more
  likely to see things distant from the transect than those near the
  transect. If this occurs, it usually indicates an issue with field
  procedure (that the distribution of objects with respect to the line,
  \(\pi(y)\) is not what we expect), so we do not want the detection
  function to model this.
\item
  \emph{Model robust}: models should be flexible enough to make many
  different shapes.
\item
  \emph{Pooling robust}: many factors can affect the probability of
  detection and it is not possible to measure all of these. We would
  like models to produce unbiased results without inclusion of these
  factors.
\item
  \emph{Estimator efficiency}: we would like models to have low
  variances, but only given they satisfy the other properties above
  (which, if they are satisfied, would give low bias).
\end{itemize}

Given these criteria, we can formulate models for \(g\).

\subsection{Formulations}\label{formulations}

There is a wide literature on possible formulations for the detection
function (Buckland 1992; Eidous 2005; Becker and Quang 2009; Giammarino
and Quatto 2014; Miller and Thomas 2015; Becker and Christ 2015).
\code{Distance} includes the most popular of these models. Here we will
detail the most popular detection function approach: ``key function plus
adjustments'' (K+A).

\subsubsection{Key function plus adjustments
(K+A)}\label{key-function-plus-adjustments-ka}

Key function plus adjustment terms (or adjustment series) models are
formulated by taking a ``key'' function and optionally adding
``adjustments'' to it to improve the fit (Buckland 1992). Mathematically
we formulate this as: \[
g(y; \boldsymbol{\theta}) = k(y; \boldsymbol{\theta}_\text{key})\left( 1+ \alpha_O(y; \boldsymbol{\theta}_\text{adjust})\right),
\] where \(k\) is the key function and \(\alpha_O\) is sum series of
functions (given in Table \ref{tab:keyadj}), described as an
\emph{adjustment of order \(O\)}. Subscripts on the parameter vector
indicate those parameters belonging to each part of the model (i.e.
\(\boldsymbol{\theta} = (\boldsymbol{\theta}_\text{key}, \boldsymbol{\theta}_\text{adjust})\)).

Available models for the key are as follows: \[
k(y) = \left\{
\begin{array}{l l}
  \exp\left(-\frac{y^2}{2 \sigma^2}\right) & \quad \text{half-normal,} \\
  1-\exp\left(\left(-\frac{y}{\sigma}\right)^{-b}\right) & \quad \text{hazard-rate,} \\
  1/w & \quad \text{uniform.}
\end{array} \right.
\] Possible modelling options for key and adjustments are given in Table
\ref{tab:keyadj} and illustrated in Figure \ref{fig:hnhr} and
\ref{fig:keyadj}. We select the number of adjustment terms (\(K\)) by
AIC (further details in ``Model checking and model selection'').

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/hn-hr-par-comp-1}

}

\caption{Half-normal (top row) and hazard-rate (bottom row) detection functions without adjustments, varying scale ($\sigma$) and (for hazard-rate) shape ($b$) parameters (values are given above the plots). On the top row from left to right, the study species becomes more detectable (higher probability of detection at larger distances). The bottom row shows the hazard-rate model's more pronounced shoulder.\label{fig:hnhr}}\label{fig:hn-hr-par-comp}
\end{figure}
\end{CodeChunk}

\begin{table}
\caption{Modelling options for key plus adjustment series models for the detection function.}
\begin{tabular}{llll}
\hline
Key function   & Form   & Adjustment series & Form\\
\hline
 Uniform  & $1/w$   & cosine  & $\sum_{o=1}^O a_o \cos(o \pi y/w)$ \\
 & & Simple polynomial & $\sum_{o=1}^O a_o (y/w)^{2o}$ \\
 Half-normal  & $\exp\left(-\frac{y^2}{2 \sigma^2}\right)$ & cosine  & $\sum_{o=2}^O a_o \cos(o \pi y/w)$ \\
 & & Hermite polynomial & $\sum_{o=2}^O a_o H_{2o}(y/\sigma)$ \\
 Hazard-rate  & $1-\exp\left[-\left(\frac{y}{\sigma}\right)^{-b}\right]$ & cosine  & $\sum_{o=2}^O a_o \cos(o \pi y/w)$ \\
 & & Simple polynomial & $\sum_{o=2}^O a_o (y/w)^{2o}$ \\
\hline
\end{tabular}
\label{tab:keyadj}
\end{table}

When adjustment terms are used it may be necessary to standardise the
results to ensure that \(g(0)=1\), so we can redefine the detection
function as: \[
g(y; \boldsymbol{\theta}) = \frac{k(y; \boldsymbol{\theta}_\text{key})\left( 1+ \alpha_O(y; \boldsymbol{\theta}_\text{adjust})\right)}{k(0; \boldsymbol{\theta}_\text{key})\left( 1+ \alpha_O(0; \boldsymbol{\theta}_\text{adjust})\right)}.
\]

A disadvantage of K+A models is that we must resort to constrained
optimisation (via the \pkg{Rsolnp} package) to ensure that the resulting
detection function is monotonic non-increasing over the whole range.

We do not always include adjustments (except in the case of the uniform
key), in which case we refer to ``key only'' models (see the next
section and ``Model checking and model selection'').

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/adjust-mix-comp-1}

}

\caption{Possible shapes for the detection function when adjustments are included for half-normal and hazard-rate models.\label{fig:keyadj}}\label{fig:adjust-mix-comp}
\end{figure}
\end{CodeChunk}

\subsubsection{Covariates}\label{covariates}

There are many factors that can affect the probability of detecting an
object. These include things like the observer, the vessel or platform
used, the sea state, weather conditions and time of day. We assume that
these variables affect detection only via the scale of the detection
function (and do not affect the shape).

Covariates can be included in this formulation by considering the scale
parameter from the half-normal or hazard-rate detection functions as a
linear model (on the exponential scale) of the (\(J\)) covariates
(\(\mathbf{z}\); a vector of length \(J\) for each observation): \[
\sigma(\mathbf{z}) = \exp(\beta_0 + \sum_{j=1}^J \beta_j z_j).
\] In the next section we will discuss model selection.

Including covariates has an important implication for our calculation of
detectability. Because we do not know what the true distribution of the
covariates is we must calculate the probability of detection conditional
on the observed values of the covariates: \[
p(\mathbf{z_i}) = \int_0^w \pi(y) g(y, \mathbf{z_i}; \boldsymbol{\theta}) dy,
\] where \(\mathbf{z_i}\) is the vector of \(J\) covariates associated
with observation \(i\). For covariate models, we are calculating a value
of ``average'' probability of detection (average in the sense of
distance) per observation. There will be as many unique values of
\(p(\mathbf{z_i})\) as there are unique covariate combinations in our
data.

K+A models that include covariates and one or more adjustments cannot be
guaranteed to be monotonic non-increasing for all covariate
combinations, as we do not have any model for the distribution of the
covariates. For this reason, we advise against using both adjustments
and covariates in a detection function (see Miller and Thomas 2015 for
an example of when this can be problematic).

\subsection{Fitting detection functions in
R}\label{fitting-detection-functions-in-r}

The workhorse of detection function fitting in \pkg{Distance} is the
\code{ds} function. Here we show off the formulations for the detection
function we have seen above applied to the minke whale and amakihi data.

\subsubsection{Minke whale}\label{minke-whale}

We can fit a model to the minke whale data, setting the truncation at
1.5km and using the default options in \code{ds} very simply:

\begin{CodeChunk}
\begin{CodeInput}
minke_hn <- ds(minke, truncation=1.5)
\end{CodeInput}
\begin{CodeOutput}
Starting AIC adjustment term selection.
Fitting half-normal key function
Key only models do not require monotonicity contraints. Not constraining model for monotonicity.
AIC= 46.872
Fitting half-normal key function with cosine(2) adjustments
AIC= 48.872

half-normal key function selected!
\end{CodeOutput}
\end{CodeChunk}

Note that \code{ds} will automatically select adjustment terms by AIC
and shows the selection steps.

Figure \ref{fig:minkeamakihi} (top left plot) shows the result of
calling \code{plot} on the resulting model object. We can also call
\code{summary} on the model object to get summary information about the
fitted model (we postpone this to the next section).

We can specify the form of the detection function via the \code{key=}
argument to \code{ds}. For example, a hazard rate model can be fitted
as:

\begin{CodeChunk}
\begin{CodeInput}
minke_hrcos <- ds(minke, truncation=1.5, key="hr")
\end{CodeInput}
\begin{CodeOutput}
Starting AIC adjustment term selection.
Fitting hazard-rate key function
Key only models do not require monotonicity contraints. Not constraining model for monotonicity.
AIC= 48.637
Fitting hazard-rate key function with cosine(2) adjustments
AIC= 50.386

hazard-rate key function selected!
\end{CodeOutput}
\end{CodeChunk}

\code{ds} fits a model with cosine adjustments (the default) but finds
the AIC improvement to be insufficient to select the adjustment.

\subsubsection{Amakihi}\label{amakihi-1}

By default \code{ds} assumes the data given to it is line transect, but
we can switch to points using \code{transect="point"}. Including
covariates in the scale is via the \code{formula=~...} argument to
\code{ds}; a hazard-rate model for the amakihi that includes observer as
a covariate can be specified by (truncating sightings at 82.5m, given in
Marques et al. 2007):

\begin{CodeChunk}
\begin{CodeInput}
amakihi_hr_obs <- ds(amakihi, truncation=82.5, transect="point",
                     key="hr", formula=~obs)
\end{CodeInput}
\begin{CodeOutput}
Cannot perform AIC adjustment term selection when covariates are used.
Fitting hazard-rate key function
AIC= 10778.448
No survey area information supplied, only estimating detection function.
\end{CodeOutput}
\end{CodeChunk}

As with the minke whale model, we can plot the resulting detection
function (Figure \ref{fig:minkeamakihi}, centre column). Because the
amakihi model included covariates, the plot shows the detection function
averaged over levels/values of the covariate. Points on the plot
indicate probability of detection for each observation. For the
\texttt{amakihi\_hr\_obs} model we see fairly clear levels of the
observer covariate in the points. Looking at the left panel of Figure
\ref{fig:minkeamakihi}, we can see this is less clear when adding
minutes after sunrise as a covariate into the model:

\begin{CodeChunk}
\begin{CodeInput}
amakihi_hr_obs_mas <- ds(amakihi, truncation=82.5, transect="point",
                         key="hr", formula=~obs+mas)
\end{CodeInput}
\begin{CodeOutput}
Cannot perform AIC adjustment term selection when covariates are used.
Fitting hazard-rate key function
AIC= 10777.376
No survey area information supplied, only estimating detection function.
\end{CodeOutput}
\end{CodeChunk}

The bottom row of Figure \ref{fig:minkeamakihi} shows the probability
density functions for the amakihi models. Probability density function
plots give a better sense of fit of the model than the detection
functions, as for point data we must rescale the histogram when plotting
the detection function to take into account the geometry of the point
sampler.

** ER suggests the g(x) plots for the amakihi be deleted **

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/minke-amakihi-hn-plot-1}

}

\caption{Top row plots of fitted detection functions overlayed on the histograms of observed distances; left minke whale data, half-normal model; centre, amakihi data hazard-rate with observer as a covariate; right, amakihi data, hazard-rate model with observer and minutes after sunrise as covariates (for amakihi histograms are rescale to take into account sampler geometry). Bottom row shows plots of the probability density function for the amakihi models (as top row). For covariate models (centre and right columns) points indicate probability of detection for a given observation (given that observations covariate values) and lines indicate the average detection function.\label{fig:minkeamakihi}}\label{fig:minke-amakihi-hn-plot}
\end{figure}
\end{CodeChunk}

\section{Model checking and model
selection}\label{model-checking-and-model-selection}

As with models fitted using \code{lm} or \code{glm} in \proglang{R}, we
can use \code{summary} to give useful information about our fitted
model. For example for our hazard-rate model for the amakihi, with
observer as a covariate:

\begin{CodeChunk}
\begin{CodeInput}
summary(amakihi_hr_obs)
\end{CodeInput}
\begin{CodeOutput}

Summary for distance analysis 
Number of observations :  1243 
Distance range         :  0  -  82.5

Model : Hazard-rate key function 
AIC   : 10778.45

Detection function parameters
Scale coefficient(s):  
              estimate         se
(Intercept) 3.06441705 0.10878121
obsTJS      0.53017364 0.09956539
obsTKP      0.08885471 0.18071851

Shape coefficient(s):  
             estimate         se
(Intercept) 0.8690009 0.06261764

                        Estimate          SE         CV
Average p              0.3142723   0.0204413 0.06504326
N in covered region 3955.1685709 274.2284029 0.06933419
\end{CodeOutput}
\end{CodeChunk}

This summary information includes details of the data and model
specification, as well as the values of the coefficients (\(\beta_j\))
and their uncertainties, an ``average'' value for the detectability (see
``Estimating abundance and variance'' for details on how this is
calculated) and its uncertainty. The final line gives an estimate of
abundance for the area covered by the survey (covered in the next
section).

\subsection{Goodness of fit}\label{goodness-of-fit}

To judge goodness of fit for detection functions when exact distances
are used, we want to compare the cumulative distribution function (CDF)
and empirical distribution function (EDF) for the detection function via
a quantile-quantile plot (Q-Q plot). In our case the CDF evaluates the
probability of observing an object at a distance less than or equal to
some given value. The EDF gives the proportion of observations for which
the CDF is less than or equal to that of a given distance. This can be
interpreted as assessing whether the number of observations up to a
given distance in line with what the model says they should be (where
`given values' are the observed distances). As usual for Q-Q plots,
``good'' models will have values close to the line \(y=x\), poor models
will show more deviations from that line.

We can inspect Q-Q plots visually, though this is prone to subjective
judgments. Instead we can quantify the Q-Q plot's information using a
Kolmogorov-Smirnov or Cramer-von Mises test (Burnham et al. 2004). Both
test whether points from the EDF and CDF are from the same distribution.
The Kolmogorov-Smirnov uses the test statistic of the largest difference
between a point on the Q-Q plot and the line \(y=x\), whereas the
Cramer-von Mises test uses the sum of all the differences. As it takes
into account more information and is therefore more powerful, the
Cramer-von Mises is generally preferred. A significant result from
either tests indicates that the EDF and CDF do not come from the same
distribution (and therefore the model is does not fit the data well).

We can generate a Q-Q plot and test results using the \code{gof_ddf}
function. We can see the goodness of fit tests below for two models for
the amakihi data (Figure \ref{amakihi-qq}), first fitting a half-normal
model without covariates or adjustments (not setting
\code{adjustment=NULL} will force \code{ds} to fit a model with no
adjustments), then calculating goodness of fit for that model and our
hazard-rate model with observer and minutes after sunrise included:

\begin{CodeChunk}
\begin{CodeInput}
amakihi_hr <- ds(amakihi, truncation=82.5, transect="point", key="hn", adjustment=NULL)
\end{CodeInput}
\begin{CodeOutput}
Fitting half-normal key function
Key only models do not require monotonicity contraints. Not constraining model for monotonicity.
AIC= 10833.841
No survey area information supplied, only estimating detection function.
\end{CodeOutput}
\begin{CodeInput}
gof_ds(amakihi_hr)
\end{CodeInput}
\begin{CodeOutput}

Goodness of fit results for ddf object

Distance sampling Kolmogorov-Smirnov test
Test statistic =  0.059345  P =  0.00031527

Distance sampling Cramer-von Mises test (unweighted)
Test statistic =  0.93083  P =  0.003578 
\end{CodeOutput}
\begin{CodeInput}
gof_ds(amakihi_hr_obs_mas)
\end{CodeInput}
\begin{CodeOutput}

Goodness of fit results for ddf object

Distance sampling Kolmogorov-Smirnov test
Test statistic =  0.036251  P =  0.076237

Distance sampling Cramer-von Mises test (unweighted)
Test statistic =  0.15016  P =  0.38908 
\end{CodeOutput}
\end{CodeChunk}

The corresponding Q-Q plots are shown in Figure \ref{amakihi-qq}.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/amakihi-qq-comp-1}

}

\caption[Comparison of quantile-quantile plots for a half-normal model (no adjustments, no covariates) and hazard-rate model with observer and minutes after sunrise for the amakihi data]{Comparison of quantile-quantile plots for a half-normal model (no adjustments, no covariates) and hazard-rate model with observer and minutes after sunrise for the amakihi data.\label{amakihi-qq}}\label{fig:amakihi-qq-comp}
\end{figure}
\end{CodeChunk}

\subsection{Model selection}\label{model-selection}

Once we have a set of models which fit well, we can use Akaike's
Information Criterion (AIC) to select between models. \pkg{Distance}
includes a function to create table of summary information for fitted
models, making it easy to get an overview of a large number of models at
once. The \code{summarize_ds_models} function takes models as input and
can be especially useful when paired with \pkg{knitr}'s \code{kable}
function to create summary tables for publication. An example of this
output is shown in Table \ref{tab:amakihi} and was generated by:

\begin{CodeChunk}
\begin{CodeInput}
library(knitr)
summarize_ds_models(amakihi_hr, amakihi_hr_obs, amakihi_hr_obs_mas)
\end{CodeInput}
\end{CodeChunk}

\begin{table}

\caption{Summary for the detection function models fitted to the amakihi data. ``C-vM'' stands for Cramer-von Mises, $P_a$ indicates average detectability (see ``Estimating abundance and variance''), se indicates standard error. Models are sorted according to AIC.\label{tab:amakihi}}
\begin{tabular}{llrrrrr}
\toprule
Key function & Formula & C-vM $p$-value & $\hat{P_a}$ & se($\hat{P_a}$) & AIC & $\Delta$AIC\\
\midrule
Hazard-rate key function & ~obs + mas & 0.389 & 0.319 & 0.020 & 10777.38 & 0.000\\
Hazard-rate key function & ~obs & 0.271 & 0.314 & 0.020 & 10778.45 & 1.073\\
Half-normal key function & ~1 & 0.004 & 0.351 & 0.011 & 10833.84 & 56.465\\
\bottomrule
\end{tabular}
\end{table}

\section{Estimating abundance and
variance}\label{estimating-abundance-and-variance}

Though fitting the detection function is the primary modelling step in
distance sampling, we are really interested in estimating detectability,
and from that abundance. We also wish to calculate our certainty in each
abundance estimate. This section addresses these issues mathematically
before showing how to estimate abundance and its variance in
\proglang{R}.

\subsection{Abundance}\label{abundance}

We wish to obtain the abundance in a study region, of which we have
sampled a (representative) subset. To do this we first calculate the
abundance in the area we have surveyed (the \emph{covered area}) to
obtain \(\hat{N}_\text{C}\), we can then scale this up to the full study
area by multiplying it by the ratio of covered area to study area.

First, to estimate abundance in the covered area (\(\hat{N}_\text{C}\)),
we use the estimates of detection probability (the
\(\{p_i; i=1,\ldots,n\}\), above) in a Horvitz-Thompson like estimator:

\begin{equation}
\hat{N}_\text{C} = \sum_{i=1}^n\frac{s_i}{\hat{p}_i},
\label{ht}
\end{equation}

where \(s_i\) are the sizes of the observed groups of objects, which is
equal to 1 if objects only occur singly (Borchers and Burnham 2004).
Thompson (2002) is the canonical reference to this type of estimator;
intuitively, we can think of the estimates of detectability
(\(\hat{p}_i\)) as ``inflating'' the group sizes (\(s_i\)) to account
for incomplete detection -- we then sum over all \(n\) detections to
obtain the abundance estimate. For models that do not include
covariates, \(\hat{p}_i\) is equal for all \(i\), so this is equivalent
to summing the groups and inflating that sum by the corresponding
\(\hat{p} (=\hat{p}_i \forall i)\).

Having obtained the abundance in the covered area, we can then scale-up
to the study area: \[
\hat{N} = \frac{A}{a} \hat{N}_\text{C},
\] where \(A\) is the area of the study region to extrapolate the
abundance estimate to and \(a\) is the covered area. For line transects
\(a=2wL\) (twice the truncation distance multiplied by the total length
of transects surveyed) and for points \(a=\pi w^2 K\).

\emph{have L and K been previously defined?}

We can use the Horvitz-Thompson-like estimator to calculate the
``average'' detectability for models which include covariates. We can
consider what single detectability value would give the estimated
\(\hat{N}\) and therefore calculate: \[
\hat{P_a} = n/\hat{N}_\text{C}.
\] This can be a useful summary statistic and is included in the
\code{summary} output and the table produced by
\code{summarize_ds_models}. \emph{anything else to say about this, seems
to mess up folks, including me}

\subsubsection{Stratification}\label{stratification}

We may wish to calculate abundance estimates for some sub-regions of the
study region, we call these areas \emph{strata}. For example, strata may
be defined by habitat types or animal gender (or some combination) which
may be of interest for biological or management reasons. To calculate
estimates for a given stratification each observation must occur in a
stratum which must be labelled with a \code{Region.Label} and have a
corresponding \code{Area} (if we are using an animal characteristic like
gender, we would have the areas be the same but if we were using say
forested vs.~wetland habitat the areas of those strata would be
different). Finally, we must also know the stratum in which transects
lie.

As an example the minke whale data consists of two strata: \code{North}
and \code{South} relating to a stratum further away and nearer the
Antarctic ice edge, respectively. Figure \ref{fig:minke-strata} shows
the two strata, along with observations and transect lines. \emph{figure
number messed up}

\begin{figure}
\includegraphics{minke-strata}
\label{fig:minke-strata}
\caption{Strata used for the minke whale data adapted from @Hedley:2004et. Points show the locations of observations along transect lines. The stepped line shows the boundary between North and South strata. Further details on the survey are available in @Branch:2001ua (simulated data is based on ``1992/93 Area III'' therein).}
\end{figure}

\subsection{Variance}\label{variance}

Here we take an intuitive approach to uncertainty estimation, for a full
derivations consult F. F. C. Marques and Buckland (2003). Uncertainty in
\(\hat{N}\) comes from two sources:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \emph{Model uncertainty}, from the estimation of the detection
  function parameters \(\boldsymbol{\theta}\)
\item
  \emph{Sampling uncertainty}, from the distribution of objects along
  the transect lines or between visiting occasions for points.
\end{enumerate}

We can see this by looking at the Horvitz-Thompson estimation in
(\ref{ht}) and consider the terms which are random. These are: the
detectability \(\hat{p}_i\) (and hence the parameters of the detection
function it is derived from) and \(n\), the number of observations. We
assume that the observed group size (\(s_i\)) is recorded without error.

Model uncertainty can be addressed using standard maximum likelihood
theory. We can invert the Hessian matrix of the likelihood to obtain a
variance-covariance matrix. We can then pre- and post-multiply this by
the derivatives of \(\hat{N}_\text{C}\) with respect to the parameters
of the detection function \[
\widehat{\text{Var}}_\text{model}\left( \hat{N}_\text{C}\right) = \left(\frac{\partial \hat{N}_\text{C}}{\partial\hat{\boldsymbol{\theta}}}\right)^\text{T} \left(\hat{\mathbf{H}}(\hat{\boldsymbol{\theta}})^{-1} \right)\frac{\partial \hat{N}_\text{C}}{\partial\hat{\boldsymbol{\theta}}}
\] where the partial derivatives of \(\hat{N}_\text{C}\) are evaluated
at the MLE (\(\hat{\theta}\)) and \(\mathbf{H}\) is the first partial
Hessian (outer product of first derivatives of the log likelihood) for
numerical stability (Buckland et al. 2001, p 62). Note that although we
calculate uncertainty in \(\hat{N}_\text{C}\), we can trivially scale-up
to variance of \(\hat{N}\).

Sampling uncertainty can be characterised (rather than as simply \(n\))
by the \emph{encounter rate}: the number of objects per unit transect.
When covariates are not included in the detection function we can define
the encounter rate as \(n/L\) for line transects (where \(L\) is the
total line length) or \(n/T\) for point transects (where \(T\) is the
total number of visits to all points) \emph{earlier eqn uses notation
\(K\) for number of points}. When covariates are included in the
detection function, it is recommended that we substitute the \(n\) in
the previous expressions with the estimated abundance \(\hat{N}\) as
this will take into account the effects of the covariates.

For line transects, by default, \pkg{Distance} uses a variation of the
estimator ``R2'' from Fewster et al. (2009) replacing number of
observations per sample with the estimated abundance per sample (Innes
et al. 2002; F. F. C. Marques and Buckland 2003): \[
\widehat{\text{Var}}_{\text{encounter},R2} = \frac{K}{L^2(K-1)} \sum_{k=1}^{K} l_k^2 \left( \frac{N_k}{l_k} - \frac{N}{L}\right)^2,
\] where \(l_k\) are the lengths of the \(K\) transects (such that
\(L = \sum_{k=1}^K l_k\)). Whereas for points we use estimator ``P3''
from Fewster et al. (2009) but again replacing \(n\) by \(\hat{N}\) to
obtain the following estimator: \[
\widehat{\text{Var}}_{\text{encounter},P3} = \frac{1}{T(K-1)} \sum_{k=1}^{K} t_k \left( \frac{N_k}{t_k} - \frac{N}{T}\right)^2,
\] where \(t_k\) is the number of visits to point \(k\) and
\(T = \sum_{k=1}^K t_k\). \emph{this messes with the \(K\) notation even
further}

Various formulations for the encounter rate variance are discussed in
detail in Fewster et al. (2009). \pkg{Distance} implements all of the
estimators of encounter rate variance given in that article. The
\code{varn} manual page gives further advice and technical detail on
encounter rate variance.

We combine detection function variance with encounter rate variance by
use of the \(\delta\)-method approximation (Goodman 1960).

\subsection{Estimating abundance and variance in
R}\label{estimating-abundance-and-variance-in-r}

Returning to the minke whale data, we have the necessary information to
calculate \(A\) and \(a\) above, so we can estimate abundance and its
variance. When we supply data to \code{ds} in the ``flatfile'' format
given above, \code{ds} will automatically calculate abundance estimates
based on the survey information in the data.

Having already fitted a model to the minke whale data, we can see the
results of the abundance estimation by viewing the model summary:

\begin{CodeChunk}
\begin{CodeInput}
summary(minke_hn)
\end{CodeInput}
\begin{CodeOutput}

Summary for distance analysis 
Number of observations :  88 
Distance range         :  0  -  1.5

Model : Half-normal key function 
AIC   : 46.87216

Detection function parameters
Scale coefficient(s):  
              estimate        se
(Intercept) -0.3411766 0.1070304

                       Estimate          SE         CV
Average p             0.5733038  0.04980421 0.08687229
N in covered region 153.4962706 17.08959835 0.11133559

Summary statistics:
  Region   Area CoveredArea  Effort  n  k         ER      se.ER     cv.ER
1  North 630582     4075.14 1358.38 49 12 0.03607238 0.01317937 0.3653591
2  South  84734     1453.23  484.41 39 13 0.08051031 0.01809954 0.2248102
3  Total 715316     5528.37 1842.79 88 25 0.04775368 0.01129627 0.2365529

Abundance:
  Label Estimate        se        cv      lcl      ucl       df
1 North 13225.44 4966.7495 0.3755450 6005.590 29124.93 12.27398
2 South  3966.46  955.9616 0.2410113 2395.606  6567.36 15.80275
3 Total 17191.90 5135.5862 0.2987212 9183.475 32184.07 14.00459

Density:
  Label   Estimate          se        cv         lcl        ucl       df
1 North 0.02097339 0.007876453 0.3755450 0.009523884 0.04618738 12.27398
2 South 0.04681073 0.011281913 0.2410113 0.028272077 0.07750560 15.80275
3 Total 0.02403400 0.007179465 0.2987212 0.012838347 0.04499280 14.00459
\end{CodeOutput}
\end{CodeChunk}

This prints a rather large amount of information. The summary can be
more concisely expressed by extracting information from the summary
object. This object is a \code{list} of \code{data.frames}, so we can
again use \pkg{knitr}s \code{kable} function to create summary tables of
abundance estimates and measures of precision, such as Table
\ref{minke-abund}.

\begin{CodeChunk}
\begin{CodeInput}
minke_table <- summary(minke_hn)$dht$individuals$N
minke_table$lcl <- minke_table$ucl <- minke_table$df <- NULL
colnames(minke_table) <- c("Stratum", "$\\hat{N}$", "$\\text{se}(\\hat{N}$)", 
                           "$\\text{CV}(\\hat{N}$)")
kable(minke_table, digits=3, format="latex", booktabs=TRUE,
      row.names=FALSE, escape=FALSE,caption="Summary of abundance estimation for the half-normal model for the minke whale data.\\label{minke-abund}")
\end{CodeInput}
\begin{table}

\caption{Summary of abundance estimation for the half-normal model for the minke whale data.\label{minke-abund}}
\begin{tabular}{lrrr}
\toprule
Stratum & $\hat{N}$ & $\text{se}(\hat{N}$) & $\text{CV}(\hat{N}$)\\
\midrule
North & 13225.44 & 4966.750 & 0.376\\
South & 3966.46 & 955.962 & 0.241\\
Total & 17191.90 & 5135.586 & 0.299\\
\bottomrule
\end{tabular}
\end{table}

\end{CodeChunk}

\section{Extensions}\label{extensions}

The features of \pkg{Distance} are deliberately limited to provide a
simplified interface for users. For more complex analyses of distance
sampling data, there are further related packages for modelling in
\proglang{R}.

We noted at the start of the article that \pkg{Distance} is a simpler
wrapper around the package \pkg{mrds}. Additional features are available
in \pkg{mrds} including the ability to model data where the assumption
that detection is certain at zero distance from the line or point by
using mark-recapture type methods when two observers are used (see Burt
et al. 2014 for an introduction).

The abundance estimates calculated here are based on the assumption that
within a given stratum density is uniform. We may extend this approach
to many strata, making the area of each very small to account for
small-scale variation in space. A more rigourous approach is to build a
spatial model incorporating spatially-referenced environmental data (for
example derived from GIS products). \pkg{Distance} interfaces with once
such package to perform this type of analysis: \pkg{dsm}. So-called
``density surface modelling'' uses the generalized additive model
framework (e.g. Wood 2006) to build models of abundance (adjusting
counts for imperfect detectability) as a function of environmental
covariates, as part of a two stage model (Hedley and Buckland 2004;
Miller et al. 2013).

\section{Conclusion}\label{conclusion}

We have given an introduction as to how to perform a distance sampling
analysis in \proglang{R}. We have covered the possible models for
detectability, model checking and selection and finally abundance and
variance estimation.

In combination with tools such as \pkg{knitr} and \pkg{rmarkdown}, the
helper functions in \pkg{Distance} provide a useful set of tools to
perform reproducible analyses of wildlife abundance for both managers
and ecologists. We hope that this paper provides useful examples for
those wishing to pursue this.

We note that there are other packages available for performing distance
sampling analyses in \proglang{R} but believe that \pkg{Distance} is the
most flexible and feature-complete. \textbf{Appendix ???} gives a
feature comparison between \pkg{Distance} and other \proglang{R}
packages for analysis of distance sampling data.

\section{Acknowledgements}\label{acknowledgements}

The authors would like to thank the many users of \pkg{Distance},
\pkg{mrds} and DISTANCE who have contributed bug reports and suggestions
for improvements over the years.

\section*{Bibliography}\label{bibliography}
\addcontentsline{toc}{section}{Bibliography}

Becker, Earl F, and Aaron M Christ. 2015. ``A Unimodal Model for Double
Observer Distance Sampling Surveys.'' \emph{PLoS ONE} 10 (8):
e0136403--18.

Becker, Earl F, and P X Quang. 2009. ``A gamma-shaped detection function
for line-transect surveys with mark-recapture and covariate data.''
\emph{Journal of Agricultural, Biological, and Environmental Statistics}
14 (2): 207--23.

Borchers, David L, and Kenneth P Burnham. 2004. ``General formulation
for distance sampling.'' In \emph{Advanced Distance Sampling}, 6--30.
Oxford University Press, Oxford, UK.

Branch, T A, and D S Butterworth. 2001. ``Southern Hemisphere minke
whales: standardised abundance estimates from the 1978/79 to 1997/98
IDCR-SOWER surveys.'' \emph{Journal of Cetacean Research and
Management}.

Buckland, Stephen T. 1992. ``Fitting Density Functions with
Polynomials.'' \emph{Applied Statistics} 41 (1): 63.

Buckland, Stephen T, David R Anderson, Kenneth P Burnham, David L
Borchers, and Len Thomas. 2001. \emph{Introduction to Distance
Sampling}. Estimating Abundance of Biological Populations. Oxford
University Press, Oxford, UK.

Buckland, Stephen T, David R Anderson, Kenneth P Burnham, Jeffrey L
Laake, David L Borchers, and Len Thomas. 2004. \emph{Advanced Distance
Sampling}. Estimating Abundance of Biological Populations. Oxford
University Press, Oxford, UK.

Buckland, Stephen T, E A Rexstad, Tiago A Marques, and Cornelia S
Oedekoven. 2015. \emph{Distance Sampling: Methods and Applications}.
Methods in Statistical Ecology. Springer International Publishing.

Burnham, Kenneth P, Stephen T Buckland, Jeffrey L Laake, David L
Borchers, Jon R B Bishop, and Len Thomas. 2004. ``Further topics in
distance sampling.'' In \emph{Advanced Distance Sampling}, 385--89.
Oxford University Press, Oxford, UK.

Burt, Mary Louise, David L Borchers, Kurt J Jenkins, and Tiago A
Marques. 2014. ``Using mark-recapture distance sampling methods on line
transect surveys.'' \emph{Methods in Ecology and Evolution} 5 (11):
1180--91.

Eidous, Omar M. 2005. ``On Improving Kernel Estimators Using Line
Transect Sampling.'' \emph{Communications in Statistics - Theory and
Methods} 34 (4): 931--41.

Fewster, Rachel M, Stephen T Buckland, Kenneth P Burnham, David L
Borchers, Peter E Jupp, Jeffrey L Laake, and Len Thomas. 2009.
``Estimating the Encounter Rate Variance in Distance Sampling.''
\emph{Biometrics} 65 (1): 225--36.

Giammarino, Mauro, and Piero Quatto. 2014. ``On estimating Hooded crow
density from line transect data through exponential mixture models.''
\emph{Environmental and Ecological Statistics} 21 (4): 689--96.

Goodman, Leo A. 1960. ``On the Exact Variance of Products.''
\emph{Journal of the American Statistical Association} 55 (292): 708.

Hedley, Sharon L, and Stephen T Buckland. 2004. ``Spatial models for
line transect sampling.'' \emph{Journal of Agricultural, Biological, and
Environmental Statistics} 9 (2): 181--99.

Innes, Stuart, M P Heide-J{ø}rgensen, Jeffrey L Laake, Kristin L Laidre,
Holly J Cleator, Pierre Richard, and Robert EA Stewart. 2002. ``Surveys
of belugas and narwhals in the Canadian High Arctic in 1996.''
\emph{NAMMCO Scientific Publications} 4 (0): 169--90.

Lahoz-Monfort, Jos{é} J, Gurutzeta Guillera-Arroita, and Brendan A
Wintle. 2013. ``Imperfect detection impacts the performance of species
distribution models.'' \emph{Global Ecology and Biogeography} 23 (4):
504--15.

Marques, Fernanda F C, and Stephen T Buckland. 2003. ``Incorporating
covariates into standard line transect analyses.'' \emph{Biometrics} 59
(4): 924--35.

Marques, Tiago A, Len Thomas, Steven G Fancy, and Stephen T Buckland.
2007. ``Improving estimates of bird density using multiple-covariate
distance sampling.'' \emph{The Auk} 124 (4): 1229.

Miller, David L, and Len Thomas. 2015. ``Mixture models for distance
sampling detection functions.'' \emph{PLoS ONE}.

Miller, David L, Mary Louise Burt, Eric A Rexstad, and Len Thomas. 2013.
``Spatial models for distance sampling data: recent developments and
future directions.'' \emph{Methods in Ecology and Evolution} 4 (11):
1001--10.

Thomas, Len, Stephen T Buckland, Eric A Rexstad, Jeffrey L Laake,
Samantha Strindberg, Sharon L Hedley, Jon R B Bishop, Tiago A Marques,
and Kenneth P Burnham. 2010. ``Distance software: design and analysis of
distance sampling surveys for estimating population size.''
\emph{Journal of Applied Ecology} 47 (1): 5--14.

Thompson, S K. 2002. \emph{Sampling}. 2nd ed. Wiley.

Wood, Simon N. 2006. \emph{Generalized Additive Models}. An Introduction
with R. CRC Press.

\end{document}