### 2016-03-31 paper.tex
stop ignoring the paper.tex file now

\documentclass[article]{jss}
\usepackage[utf8]{inputenc}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\author{
David L Miller\\Integrated Statistics and University of St Andrews \AND Eric Rexstad\\University of St Andrews \And Len Thomas\\University of St Andrews \AND Laura Marshall\\University of St Andrews \And Jeffrey L Laake\\Marine Mammal Laboratory
}
\title{Distance Sampling in \proglang{R}}
\Keywords{distance sampling, abundance estimation, line transect, point transect, detection function, Horvitz-Thompson, \proglang{R}, Distance}

\Abstract{
Estimating the abundance and spatial distribution of animal and plant
populations is essential for conservation and management. We introduce
the \proglang{R} package \pkg{Distance} that implements distance
sampling methods to estimate abundance. We describe how users can obtain
estimates of abundance (and density) using the package as well
documenting the links it provides with other more specialized
\proglang{R} packages. We also demonstrate how \pkg{Distance} provides a
migration pathway from previous software, thereby allowing us to deliver
cutting-edge methods to the users more quickly.
}

\Plainauthor{David L Miller, Eric Rexstad, Len Thomas, Laura Marshall, Jeffrey L Laake}
\Plaintitle{Distance Sampling in R}
\Plainkeywords{distance sampling, abundance estimation, line transect, point transecs, detection function, Horvitz-Thompson, R, Distance}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
\Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
    David L Miller\\
  Integrated Statistics and University of St Andrews\\
  16 Sumner Street, Woods Hole, MA 02543, USA and Centre for Research into
  Ecological and Environmental Modelling, The Observatory, St Andrews,
  Fife KY16 9LZ, Scotland\\
  E-mail: \href{mailto:dave@ninepointeightone.net}{\nolinkurl{dave@ninepointeightone.net}}\\
  URL: \url{http://converged.yt}\\~\\
      Eric Rexstad\\
  University of St Andrews\\
  Centre for Research into Ecological and Environmental Modelling, The
  Observatory, St Andrews, Fife KY16 9LZ, Scotland\\
  E-mail: \href{mailto:Eric.Rexstad@st-andrews.ac.uk}{\nolinkurl{Eric.Rexstad@st-andrews.ac.uk}}\\

      Len Thomas\\
  University of St Andrews\\
  Centre for Research into Ecological and Environmental Modelling, The
  Observatory, St Andrews, Fife KY16 9LZ, Scotland\\
  E-mail: \href{mailto:len.thomas@st-andrews.ac.uk}{\nolinkurl{len.thomas@st-andrews.ac.uk}}\\
  URL: \url{http://lenthomas.org}\\~\\
      Laura Marshall\\
  University of St Andrews\\
  Centre for Research into Ecological and Environmental Modelling, The
  Observatory, St Andrews, Fife KY16 9LZ, Scotland\\
  E-mail: \href{mailto:lhm@st-andrews.ac.uk}{\nolinkurl{lhm@st-andrews.ac.uk}}\\

      Jeffrey L Laake\\
  Marine Mammal Laboratory\\
  Alaska Fisheries Science Center National Oceanic and Atmospheric
  Administration 7600 Sand Point Way N.E., Seattle, WA 98115, USA\\
  E-mail: \href{mailto:Jeff.Laake@noaa.gov}{\nolinkurl{Jeff.Laake@noaa.gov}}\\

  }

\usepackage{amsmath} \usepackage{amssymb} \usepackage{longtable}
\usepackage{booktabs}
\usepackage[colorinlistoftodos, color=green!40, textwidth=60, textsize=small]{todonotes}

\begin{document}

\section{Introduction}\label{introduction}

Distance sampling
\citep{Buckland:2001vm, Buckland:2004ts, buckland2015distance}
encompasses a suite of methods used to estimate the density and/or
abundance of biological populations. Distance sampling can be thought of
as an extension of plot sampling. Plot sampling involves selecting a
number of plots (small areas) at random within the study area and
counting the objects of interest that are contained within each plot. By
selecting the plots at random we can assume that the density of objects
in the plots is representative of the study area as a whole. One of the
key assumptions of plot sampling is that all objects within each of the
plots are counted. Distance sampling relaxes this assumption in that
observers are no longer required to detect (either by eye, video/audio
recording etc) and count everything within selected plots. While plot
sampling techniques are adequate for static populations occurring at
high density they are inefficient for more sparsely distributed
populations. Distance sampling provides a more efficient solution in
such circumstances.

Conventional distance sampling assumes the observer is located either at
a point or moving along a line and will observe all objects that occur
at the point or on the line. The further away an object is from the
point or line (also known as the sampler or transect) the less likely it
is that the observer will see it. We can use the distances to each of
the detected objects from the line or point to build a model of the
probability of detection given distance from the sampler -- the
\emph{detection function}. The detection function can be used to infer
how many objects were missed and thereby produce estimates of density
and/or abundance.

The Windows program Distance (or ``DISTANCE''; for clarity henceforth
``Distance for Windows'') can be used to fit detection functions to
distance sampling data. It was first released (versions 1.0 - 3.0;
principally programmed by Jeff Laake while working at the National
Marine Mammal Laboratory) as a console-based application (this in turn
was based on earlier software TRANSECT, \citet{Burnham:1980wz} and
algorithms developed in \citet{Buckland:1992fa}), before the first
graphical interface (Distance for Windows 3.5) was released in November
1998. Since this time it has evolved to include various design and
analysis features \citep{Thomas:2010cf}. Distance for Windows versions 5
onwards have included \proglang{R} \citep{rcore} packages as the
analysis engines providing additional, more complex analysis options
than those offered by the original (FORTRAN) code.

As Distance for Windows becomes increasingly reliant on analyses
performed in \proglang{R} and many new methods are being developed, we
are encouraging the use of our \proglang{R} packages directly.
\proglang{R} provides a huge variety of functionality for data
exploration and reproducible research, much more than is possible in
Distance for Windows.

Until now those wishing to use our \proglang{R} packages for straight
forward distance sampling analyses would have had to negotiate the
package \pkg{mrds} \citep{mrds-pkg} designed for mark-recapture distance
sampling \citep{Burt:2014gu}, requiring a complex data structure to
perform analyses. \pkg{Distance} is a wrapper package around \pkg{mrds}
making it easier to get started with basic distance sampling analyses in
\proglang{R}. The most basic detection function estimation only requires
a numeric vector of distances. Here we demonstrate how to use
\pkg{Distance} to fit detection functions, perform model checking and
selection, and estimate abundance.

\subsection{Distance sampling}\label{distance-sampling}

The distribution of the observed distances is a product of detectability
\citep[sometimes referred to as ``perception bias'';][]{Marsh:1989ho}
and the distribution of the animals with respect to the line or point.
Our survey design allows us to assume a distribution for the animals
with respect to the sampler.

For line transect studies we assume that objects are uniformly
distributed with respect to the line (i.e., the number of animals
available for detection is the same at all distances). For point
transect surveys, area increases linearly with radial distance, implying
a triangular distribution with respect to the point. Figure
\ref{fig:detectiondist} shows how these distributions, when combined
with a detection function, give rise to the observed distribution of
recorded distances.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/detectdists-1}

}

\caption{Panels show an example detection function (left), the pdf of object distances (middle) and the resulting pdf of detection distances (right) for line transects (top row) and point transects (bottom row). The pdf of observed detection distances in the right hand plots are obtained by multiplying the detection function by the pdf of object distances and rescaling. In this example, detection probability becomes effectively zero at 500 (distances shown on $x$-axis are arbitrary). \label{fig:detectiondist}}\label{fig:detectdists}
\end{figure}
\end{CodeChunk}

Figure \ref{fig:pointslines} shows simulated sampling of a population of
500 objects using line and point transects and their corresponding
histograms of observed detection distances. Note that for the purposes
of distance sampling an ``object'' may either refer to an individual in
a population or a cluster of individuals.

\begin{CodeChunk}
\begin{CodeOutput}
Warning: package 'nlme' was built under R version 3.2.4
\end{CodeOutput}
\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/points-and-lines-1}

}

\caption{Left side plots show an example of a survey of a (1x1) area containing a population of 500 objects; blue dashed lines (top plot) and triangles (bottom plot) indicate sampler placement, red dots indicate detected individuals and grey dots give the locations of unobserved individuals. The right side of the figure shows histograms of observed distances (again, lines top and points bottom).\label{fig:pointslines}}\label{fig:points-and-lines}
\end{figure}
\end{CodeChunk}

Good survey design is essential to produce reliable density and
abundance estimates from distance sampling data. Survey design is beyond
the scope of this article but we refer readers to \citet[Chapter
7]{Buckland:2001vm} and \citet[Chapter 2]{buckland2015distance} for
introductory information; \citet{Strindberg:2004vr} contains information
on automated survey design using geographical information systems (GIS);
\citet{Thomas:2007wz} gives an example of designing a distance sampling
survey in a complex region. We also note that Distance for Windows
includes a GIS survey design tool \citep{Thomas:2010cf}.

\pkg{Distance} provides a selection of candidate functions to describe
the probability of detection and estimates the associated parameters
using maximum likelihood estimation. The probability of detecting an
object may not only depend on how far it is from the observer but also
on other factors such as weather conditions, ground cover, cluster size
etc. The \pkg{Distance} package also allows the incorporation of such
covariates into the detection function allowing the detection function
scale parameter to vary based on these covariates.

Having estimated the detection function's parameters, one can then
integrate out distance from the function (as the detection function
describes the probability of detection \emph{given} distance) to get an
``average'' probability of detection, which can be used to correct the
observed counts. Summing the corrected counts gives an estimate of
abundance in the area covered by surveys, which can be multiplied up to
the total study area.

In addition to randomly placed samplers distance sampling relies on
three other main assumptions. Firstly, all objects directly on the
transect line or point (i.e.~those at zero distance) are detected (see
``Extensions'' for methods to deal with the situation when this is not
possible). Secondly, objects are stationary or detected prior to any
movement. Thirdly, distance to the object must be measured accurately,
or the observation allocated to the correct distance bin for grouped
data. Depending on the survey species some of these assumptions may be
more difficult to meet than others. Further information on field methods
to help meet these assumptions can be found in \citet{Buckland:2001vm}
and \citet{buckland2015distance}.

The rest of the paper has the following structure: we describe data
formatting for \pkg{Distance}; candidate detection function models are
described in terms of formulation and fitted examples in \proglang{R}.
We then show how to perform model checking, goodness of fit and model
selection. We go on to show how to estimate abundance, including
stratified estimates of abundance. The final two sections of the article
look at extensions (both in terms of methodology and software) and put
the package in a broader context amongst other \proglang{R} packages
used for estimating the abundance of biological populations from
distance sampling data.

\section{Data}\label{data}

We introduce two example analyses performed in \pkg{Distance}: one line
transect and one point transect. These data sets have been chosen as
they represent typical data seen in practice. The below example analyses
are not intended to serve as guidelines, but to demonstrate features of
the software. Practical advice on approaches to analysis is given in
\citet{Thomas:2010cf}.

\subsubsection{Minke whales}\label{minke-whales}

The line transect data have been simulated from models fitted to
Antarctic minke whale (\emph{Balaenoptera bonaerensis}) data. These data
were collected as part of the International Whaling Commission's
International Decade of Cetacean Research Southern Ocean Whale and
Ecosystem Research (IWC IDCR-SOWER) programme 1992-1993 austral summer
surveys. These data consist of 99 observations that are stratified based
on location (near or distant from ice edge) and effort data (transect
lengths). Further details on the survey are available in
\citet{Branch:2001ua} (data is simulated based on the design used for
``1992/93 Area III'' therein).

\subsubsection{Amakihi}\label{amakihi}

The point transect data set consists of 1485 observations of Amakihi
(\emph{Hemignathus virens}; a Hawaiian songbird), collected at 41 points
between 1992 and 1995. The data include distances and two covariates
collected during the survey: observer (a three level factor), time after
sunrise (transformed to minutes (continuous) or hours (factor)
covariates). Data are analysed comprehensively in
\citet{Marques:2007ey}.

\subsection{Data setup}\label{data-setup}

Generally, data collected in the field will require some formatting
before use with \pkg{Distance}, though there are a range of possible
formats, dependent on the model specification and the output required:

\begin{itemize}
\tightlist
\item
  In the simplest case, where the objective is to estimate a detection
  function and exact distances are collected, all that is required is a
  numeric vector of distances.
\item
  To include additional covariates into the detection function (see
  ``Detection functions'') a \code{data.frame} is required. Each row of
  the \code{data.frame} contains the data on one observation. The
  \code{data.frame} must contain a column named \code{distance}
  (containing the observed distances) and additional named columns for
  any covariates that may affect detectability (for example
  \code{observer} or \code{seastate}). The column name \code{size} is
  reserved for the cluster sizes (sometimes referred to as group sizes)
  in this case each row represents an observation of a cluster rather
  than individual (see \citet{Buckland:2001vm}, Section 1.4.3 for more
  on defining clusters and ``Extensions'' below for one approach to
  dealing with uncertain cluster size). Additional reserved names
  include \code{object} and \code{detected}, these are not required for
  conventional distance sampling and should be avoided (see
  ``Extensions'' for an explanation of their use).
\item
  To estimate density or to estimate abundance beyond the sampled area,
  additional information is required. Additional columns should be
  included in the \code{data.frame} specifying: \code{Sample.Label}, the
  ID of the transect; \code{Effort}, transect effort (for lines their
  length and for points the number of times that point was visited);
  \code{Region.Label}, the stratum containing the transect (which may be
  from pre- or post-survey stratification, see ``Estimating abundance
  and variance''); \code{Area}, the area of the strata. Transects which
  were surveyed but have no observations must be included in the data
  set with \code{NA} for \code{distance} and any other covariates. We
  refer to this data format (where all information is contained in one
  table) as ``flatfile'' as it can be easily created in a single
  spreadsheet.
\end{itemize}

As we will see in ``Extensions'', further information is also required
for fitting more complex models.

It is also possible to use distances collected in intervals (``binned''
or ``grouped'' data, as opposed to ``exact'' distances) to model the
detection function. In this case the column \code{distance} is replaced
by two columns \code{distbegin} and \code{distend} referring to the
distance interval start and end cutpoints. More information on binned
data is included in \citet{Buckland:2001vm} sections 4.5 and 7.4.1.2.

The columns \code{distance}, \code{Area} and (in the case of line
transects) \code{Effort} have associated units (though these are not
explicitly included in a \pkg{Distance} analysis). We recommend that
data in these columns are converted to SI units before starting any
analysis to ensure that resulting abundance and density estimates have
sensible units. For example, if distances from a line transect survey
are recorded in metres, the \code{Effort} columns should contain line
lengths also in metres and the \code{Area} column gives the stratum
areas in square metres. This would lead to density estimates of animals
per square metre.

The minke whale data follows the ``flatfile'' format given in the last
bullet point:

\begin{CodeChunk}
\begin{CodeInput}
library("Distance")
head(minke)
\end{CodeInput}
\begin{CodeOutput}
  Region.Label  Area Sample.Label Effort distance
1        South 84734            1  86.75     0.10
2        South 84734            1  86.75     0.22
3        South 84734            1  86.75     0.16
4        South 84734            1  86.75     0.78
5        South 84734            1  86.75     0.21
6        South 84734            1  86.75     0.95
\end{CodeOutput}
\end{CodeChunk}

Whereas the amakihi data lacks effort and stratum data:

\begin{CodeChunk}
\begin{CodeInput}
head(amakihi)
\end{CodeInput}
\begin{CodeOutput}
   survey object distance obs mas has detected
1 July 92      1       40 TJS  50   1        1
2 July 92      2       60 TJS  50   1        1
3 July 92      3       45 TJS  50   1        1
4 July 92      4      100 TJS  50   1        1
5 July 92      5      125 TJS  50   1        1
6 July 92      6      120 TJS  50   1        1
\end{CodeOutput}
\end{CodeChunk}

We will explore the consequences of including effort and stratum data in
the analysis below.

\section{Detection functions}\label{detection-functions}

The detection function models the probability
\(\mathbb{P}(\text{object detected } \vert \text{ object at distance } y)\)
and is usually denoted \(g(y; \boldsymbol{\theta})\) where \(y\) is
distance (from a line or point) and \(\boldsymbol{\theta}\) is a vector
of parameters to be estimated. Our goal is to estimate an \emph{average
probability of detection} (\(p\), average in the sense of an average
over distance from \(0\) to truncation distance \(w\)), so we must
integrate out distance (\(y\)) from the detection function: \[
p = \int_0^w \pi(y) g(y; \boldsymbol{\theta}) dy
\] where \(\pi(y)\) describes the distribution of objects with respect
to the sampler; \(\pi(x)=1/w\) for line transects and
\(\pi(r)=\frac{2r}{w^2}\) for point transects, taking into account the
geometry of the sampler \citep[usually referred to as the
\emph{probability density function of distances};][Chapter
3]{Buckland:2001vm}. When considering a particular transect type we let
\(x\) denote a perpendicular distance from a line and \(r\) denote
radial distance from a point (rather than using \(y\)).

It is crucial that the detection function accurately models
detectability at small distances; we are less worried by its behaviour
further away from 0. To ensure that the model is not overly influenced
by distances far from zero, we truncate the distances beyond a given
distance \(w\) (this can be done during analysis or while collecting
data in the field).

Models for the detection function are expected to have the following
properties \citep[Chapter 5]{buckland2015distance}:

\begin{itemize}
\tightlist
\item
  \emph{Shoulder}: we expect observers to be able to see objects near
  them, not just those directly in front of them. For this reason, we
  expect the detection function to be flat near the line or point.
\item
  \emph{Non-increasing}: we do not think that observers should be more
  likely to see distant objects than those nearer the transect. If this
  occurs, it usually indicates an issue with survey design or field
  procedures (for example that the distribution of objects with respect
  to the line, \(\pi(y)\) is not what we expect), so we do not want the
  detection function to model this
  \citep{Marques:2010he, Marques:2012fy, Miller:2015hw}.
\item
  \emph{Model robust}: models should be flexible enough to fit many
  different shapes.
\item
  \emph{Pooling robust}: many factors can affect the probability of
  detection and it is not possible to measure all of these. We would
  like models to produce unbiased results without inclusion of these
  factors.
\item
  \emph{Estimator efficiency}: we would like models to have low
  variances, given they satisfy the other properties above (which, if
  satisfied, would give low bias).
\end{itemize}

Given these criteria, we can formulate models for \(g\).

\subsection{Formulations}\label{formulations}

There is a wide literature on possible formulations for the detection
function
\citep{Buckland:1992fa, Eidous:2005bj, Becker:2009cj, Giammarino:2014eg, Miller:2015hw, Becker:2015fi}.
\code{Distance} includes the most popular of these models. Here we
detail the most popular detection function approach: ``key function plus
adjustments'' (K+A).

\subsubsection{Key function plus adjustments
(K+A)}\label{key-function-plus-adjustments-ka}

Key function plus adjustment terms (or adjustment series) models are
formulated by taking a ``key'' function and optionally adding
``adjustments'' to it to improve the fit \citep{Buckland:1992fa}.
Mathematically we formulate this as: \[
g(y; \boldsymbol{\theta}) = k(y; \boldsymbol{\theta}_\text{key})\left( 1+ \alpha_O(y; \boldsymbol{\theta}_\text{adjust})\right),
\] where \(k\) is the key function and \(\alpha_O\) is sum series of
functions (given in Table \ref{tab:keyadj}), described as an
\emph{adjustment of order \(O\)}. Subscripts on the parameter vector
indicate those parameters belonging to each part of the model (i.e.
\(\boldsymbol{\theta} = (\boldsymbol{\theta}_\text{key}, \boldsymbol{\theta}_\text{adjust})\)).

Available models for the key are as follows: \[
k(y) = \left\{
\begin{array}{l l}
  \exp\left(-\frac{y^2}{2 \sigma^2}\right) & \quad \text{half-normal,} \\
  1-\exp\left(\left(-\frac{y}{\sigma}\right)^{-b}\right) & \quad \text{hazard-rate,} \\
  1/w & \quad \text{uniform.}
\end{array} \right.
\] Possible modelling options for key and adjustments are given in Table
\ref{tab:keyadj} and illustrated in Figures \ref{fig:hnhr} and
\ref{fig:keyadj}. We select the number of adjustment terms (\(K\)) by
AIC (further details in ``Model checking and model selection'').

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/hn-hr-par-comp-1}

}

\caption{Half-normal (top row) and hazard-rate (bottom row) detection functions without adjustments, varying scale ($\sigma$) and (for hazard-rate) shape ($b$) parameters (values are given above the plots). On the top row from left to right, the study species becomes more detectable (higher probability of detection at larger distances). The bottom row shows the hazard-rate model's more pronounced shoulder.\label{fig:hnhr}}\label{fig:hn-hr-par-comp}
\end{figure}
\end{CodeChunk}

\begin{table}
\begin{tabular}{llll}
\hline
Key function   & Form   & Adjustment series & Form\\
\hline
 Uniform  & $1/w$   & cosine  & $\sum_{o=1}^O a_o \cos(o \pi y/w)$ \\
 & & Simple polynomial & $\sum_{o=1}^O a_o (y/w)^{2o}$ \\
 Half-normal  & $\exp\left(-\frac{y^2}{2 \sigma^2}\right)$ & cosine  & $\sum_{o=2}^O a_o \cos(o \pi y/w)$ \\
 & & Hermite polynomial & $\sum_{o=2}^O a_o H_{2o}(y/\sigma)$ \\
 Hazard-rate  & $1-\exp\left[-\left(\frac{y}{\sigma}\right)^{-b}\right]$ & cosine  & $\sum_{o=2}^O a_o \cos(o \pi y/w)$ \\
 & & Simple polynomial & $\sum_{o=2}^O a_o (y/w)^{2o}$ \\
\hline
\end{tabular}
\label{tab:keyadj}
\caption{Modelling options for key plus adjustment series models for the detection function. For each key function the default adjustments are cosine in \pkg{Distance}. Note that in the adjustments functions distance is divided by the width or the scale parameter to ensure the shape of adjustment independent of the units of $y$ (Marques et al., 2007); defaults are shown here, though either can be selected to rescale the distances.}
\end{table}

When adjustment terms are used it is necessary to standardise the
results to ensure that \(g(0)=1\), so we can redefine the detection
function as: \[
g(y; \boldsymbol{\theta}) = \frac{k(y; \boldsymbol{\theta}_\text{key})\left( 1+ \alpha_O(y; \boldsymbol{\theta}_\text{adjust})\right)}{k(0; \boldsymbol{\theta}_\text{key})\left( 1+ \alpha_O(0; \boldsymbol{\theta}_\text{adjust})\right)}.
\]

A disadvantage of K+A models is that we must resort to constrained
optimisation \citep[via the \pkg{Rsolnp} package;][]{rsolnp-pkg} to
ensure that the resulting detection function is monotonic non-increasing
over its range.

It is not always necessary to include adjustments (except in the case of
the uniform key) and in such cases we refer to these as ``key only''
models (see the next section and ``Model checking and model
selection'').

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/adjust-mix-comp-1}

}

\caption{Possible shapes for the detection function when cosine adjustments are included for half-normal and hazard-rate models.\label{fig:keyadj}}\label{fig:adjust-mix-comp}
\end{figure}
\end{CodeChunk}

\subsubsection{Covariates}\label{covariates}

There are many factors that can affect the probability of detecting an
object: observer skill, cluster size (if objects occur in clusters), the
vessel or platform used, sea state, other weather conditions, time of
day and more. In \pkg{Distance} we assume that these variables affect
detection only via the scale of the detection function (and do not
affect the shape).

Covariates can be included in this formulation by considering the scale
parameter from the half-normal or hazard-rate detection functions as a(n
exponentiated) linear model of the (\(J\)) covariates (\(\mathbf{z}\); a
vector of length \(J\) for each observation): \[
\sigma(\mathbf{z}) = \exp(\beta_0 + \sum_{j=1}^J \beta_j z_j).
\]

Including covariates has an important implication for our calculation of
detectability. We do not know the true distribution of the covariates,
we therefore must either: \((i)\) model the distribution of the
covariates and integrate the covariates out of the joint density (thus
making strong assumptions about their distribution), or \((ii)\)
calculate the probability of detection conditional on the observed
values of the covariates \citep{Marques:2003vb}. We opt for the latter:
\[
p(\mathbf{z_i}) = \int_0^w \pi(y) g(y, \mathbf{z_i}; \boldsymbol{\theta}) dy,
\] where \(\mathbf{z_i}\) is the vector of \(J\) covariates associated
with observation \(i\). For covariate models, we calculate a value of
``average'' probability of detection (average in the sense of distance
being integrated out) per observation. There are as many unique values
of \(p(\mathbf{z_i})\) as there are unique covariate combinations in our
data.

K+A models that include covariates and one or more adjustments cannot be
guaranteed to be monotonic non-increasing for all covariate
combinations. Without a model for the distribution of the covariates, it
is not possible to know what the behaviour of the detection function
will be across the ranges of the covariates. As such we cannot set
meaningful constraints on monotonicity. For this reason, we advise
caution when using both adjustments and covariates in a detection
function \citep[see][ for an example of when this can be
problematic]{Miller:2015hw}.

\subsection{Fitting detection functions in
R}\label{fitting-detection-functions-in-r}

A detection function can be fitted in \pkg{Distance} using the \code{ds}
function. Here we apply some of the possible formulations for the
detection function we have seen above to the minke whale and amakihi
data.

\subsubsection{Minke whale}\label{minke-whale}

First we fit a model to the minke whale data, setting the truncation at
1.5km and using the default options in \code{ds} very simply:

\begin{CodeChunk}
\begin{CodeInput}
minke_hn <- ds(minke, truncation = 1.5)
\end{CodeInput}
\begin{CodeOutput}
Starting AIC adjustment term selection.
\end{CodeOutput}
\begin{CodeOutput}
Fitting half-normal key function
\end{CodeOutput}
\begin{CodeOutput}
Key only model: not constraining for monotonicity.
\end{CodeOutput}
\begin{CodeOutput}
AIC= 46.872
\end{CodeOutput}
\begin{CodeOutput}
Fitting half-normal key function with cosine(2) adjustments
\end{CodeOutput}
\begin{CodeOutput}
AIC= 48.872
\end{CodeOutput}
\begin{CodeOutput}

Half-normal key function selected.
\end{CodeOutput}
\end{CodeChunk}

Note that when there are no covariates in the model, \code{ds} will add
adjustment terms to the model until there is no improvement in AIC.

Figure \ref{fig:minkeamakihi} (left panel) shows the result of calling
\code{plot} on the resulting model object. We can also call
\code{summary} on the model object to get summary information about the
fitted model (we postpone this until the next section).

A different form for the detection function can be specified via the
\code{key=} argument to \code{ds}. For example, a hazard rate model can
be fitted as:

\begin{CodeChunk}
\begin{CodeInput}
minke_hrcos <- ds(minke, truncation = 1.5, key = "hr")
\end{CodeInput}
\begin{CodeOutput}
Starting AIC adjustment term selection.
\end{CodeOutput}
\begin{CodeOutput}
Fitting hazard-rate key function
\end{CodeOutput}
\begin{CodeOutput}
Key only model: not constraining for monotonicity.
\end{CodeOutput}
\begin{CodeOutput}
AIC= 48.637
\end{CodeOutput}
\begin{CodeOutput}
Fitting hazard-rate key function with cosine(2) adjustments
\end{CodeOutput}
\begin{CodeOutput}
AIC= 50.386
\end{CodeOutput}
\begin{CodeOutput}

Hazard-rate key function selected.
\end{CodeOutput}
\end{CodeChunk}

Here \code{ds} also fits the hazard-rate model then hazard-rate with a
cosine adjustment but the AIC improvement is insufficient to select the
adjustment, so the hazard-rate key-only model is returned.

Other adjustment series can be selected using the \texttt{adjustment=}
argument and specific orders of adjustments can be set using
\texttt{order=}. For example, to specify a uniform model with cosine
adjustments of order 1 and 2 we can write:

\begin{CodeChunk}
\begin{CodeInput}
minke_unifcos <- ds(minke, truncation = 1.5, key = "unif",
                    adjustment = "cos", order = c(1, 2))
\end{CodeInput}
\begin{CodeOutput}
Fitting uniform key function with cosine(1,2) adjustments
\end{CodeOutput}
\begin{CodeOutput}
AIC= 48.268
\end{CodeOutput}
\end{CodeChunk}

Hermite polynomial adjustments use the code \texttt{"herm"} and simple
polynomials \texttt{"poly"}, adjustment order should be in line with
Table \ref{tab:keyadj}.

\subsubsection{Amakihi}\label{amakihi-1}

\code{ds} assumes the data given to it has been collected as line
transects, but we can switch to point transects using the argument
\code{transect="point"}. We can include covariates in the scale
parameter via the \code{formula=~...} argument to \code{ds}. A
hazard-rate model for the amakihi that includes observer as a covariate
and a truncation distance of 82.5m \citep{Marques:2007ey} can be
specified using :

\begin{CodeChunk}
\begin{CodeInput}
amakihi_hr_obs <- ds(amakihi, truncation = 82.5, transect = "point",
                     key = "hr", formula = ~obs)
\end{CodeInput}
\begin{CodeOutput}
Model contains covariate term(s): no adjustment terms will be included.
\end{CodeOutput}
\begin{CodeOutput}
Fitting hazard-rate key function
\end{CodeOutput}
\begin{CodeOutput}
AIC= 10778.448
\end{CodeOutput}
\begin{CodeOutput}
No survey area information supplied, only estimating detection function.
\end{CodeOutput}
\end{CodeChunk}

Note that here, unlike with the minke whale data, \code{ds} warns us
that we have only supplied enough information to estimate the detection
function (not density or abundance).

While automatic AIC selection is performed on adjustment terms, model
selection for covariates must be performed manually. Here we add a
second covariate: minutes after sunrise. We will compare these two
models further in the following section.

\begin{CodeChunk}
\begin{CodeInput}
amakihi_hr_obs_mas <- ds(amakihi, truncation = 82.5, transect = "point",
                         key = "hr", formula = ~obs + mas)
\end{CodeInput}
\begin{CodeOutput}
Model contains covariate term(s): no adjustment terms will be included.
\end{CodeOutput}
\begin{CodeOutput}
Fitting hazard-rate key function
\end{CodeOutput}
\begin{CodeOutput}
AIC= 10777.376
\end{CodeOutput}
\begin{CodeOutput}
No survey area information supplied, only estimating detection function.
\end{CodeOutput}
\end{CodeChunk}

As with the minke whale model, we can plot the resulting models (Figure
\ref{fig:minkeamakihi}, middle and right panels). However, for point
transect studies, probability density function plots give a better sense
of model fit than the detection function plots. This is because when
plotting the detection function for point transect data, the histogram
must be rescaled to account for the geometry of the point sampler. The
amakihi models included covariates, so the plots show the detection
function averaged over levels/values of the covariate. Points on the
plot indicate probability of detection for each observation. For the
\texttt{amakihi\_hr\_obs} model we see fairly clear levels of the
observer covariate in the points. Looking at the right panel of Figure
\ref{fig:minkeamakihi}, this is less clear when adding minutes after
sunrise (a continuous covariates) to the model.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/minke-amakihi-hn-plot-1}

}

\caption{Left: fitted detection function overlayed on the histogram of observed distances for the minke whale data using half-normal model. Centre and right: plots of the probability density function for the amakihi models. Centre, hazard-rate with observer as a covariate; right, hazard-rate model with observer and minutes after sunrise as covariates. Points indicate probability of detection for a given observation (given that observation's covariate values) and lines indicate the average detection function.\label{fig:minkeamakihi}}\label{fig:minke-amakihi-hn-plot}
\end{figure}
\end{CodeChunk}

\section{Model checking and model
selection}\label{model-checking-and-model-selection}

As with models fitted using \code{lm} or \code{glm} in \proglang{R}, we
can use \code{summary} to give useful information about our fitted
model. For our hazard-rate model of the amakihi data, with observer as a
covariate:

\begin{CodeChunk}
\begin{CodeInput}
summary(amakihi_hr_obs)
\end{CodeInput}
\begin{CodeOutput}

Summary for distance analysis 
Number of observations :  1243 
Distance range         :  0  -  82.5

Model : Hazard-rate key function 
AIC   : 10778.45

Detection function parameters
Scale coefficient(s):  
              estimate         se
(Intercept) 3.06441705 0.10878121
obsTJS      0.53017364 0.09956539
obsTKP      0.08885471 0.18071851

Shape coefficient(s):  
             estimate         se
(Intercept) 0.8690009 0.06261764

                        Estimate          SE         CV
Average p              0.3142723   0.0204413 0.06504326
N in covered region 3955.1685709 274.2284029 0.06933419
\end{CodeOutput}
\end{CodeChunk}

This summary information includes details of the data and model
specification, as well as the values of the coefficients (\(\beta_j\))
and their uncertainties, an ``average'' value for the detectability (see
``Estimating abundance and variance'' for details on how this is
calculated) and its uncertainty. The final line gives an estimate of
abundance in the area covered by the survey (see the next section;
though note this estimate does not take into account cluster size).

\subsection{Goodness of fit}\label{goodness-of-fit}

We use a quantile-quantile plot (Q-Q plot) to visually assess how well a
detection functions fits the data when we have exact distances. Here the
Q-Q plot compares the cumulative distribution function (CDF) of the
fitted detection function to the distribution of the data (empirical
distribution function or EDF). The Q-Q plots in \pkg{Distance} plot a
point for every observation. The EDF is the proportion of points that
have been observed at a distance equal to or less than the distance of
that observation. The CDF is calculated from the fitted detection
function as the probability of observing an object at a distance less
than or equal to that of the given observation. This can be interpreted
as assessing whether the number of observations up to a given distance
is in line with what the model says they should be. As usual for Q-Q
plots, ``good'' models will have values close to the line \(y=x\), poor
models will show greater deviations from that line.

Q-Q plots can be inspected visually, though this is prone to subjective
judgments. Therefore, we also quantify the Q-Q plot's information using
a Kolmogorov-Smirnov or Cramér-von Mises test \citep{Burnham:2004vd}.
Both test whether points from the EDF and CDF are from the same
distribution. The Kolmogorov-Smirnov uses the test statistic of the
largest difference between a point on the Q-Q plot and the line \(y=x\),
whereas the Cramér-von Mises test uses the sum of all the differences.
As it takes into account more information and is therefore more
powerful, the Cramér-von Mises is generally preferred. A significant
result from either test gives evidence against the null hypothesis (that
the data arose from the fitted model), suggesting that the model does
not fit the data well.

We can generate a Q-Q plot and test results using the \code{gof_ds}
function. Figure \ref{amakihi-qq} shows the goodness of fit tests for
two models for the amakihi data. We first fit a half-normal model
without covariates or adjustments (setting \code{adjustment=NULL} will
force \code{ds} to fit a model with no adjustments):

\begin{CodeChunk}
\begin{CodeInput}
amakihi_hn <- ds(amakihi, truncation = 82.5, transect = "point",
                 key = "hn", adjustment = NULL)
\end{CodeInput}
\begin{CodeOutput}
Fitting half-normal key function
\end{CodeOutput}
\begin{CodeOutput}
Key only model: not constraining for monotonicity.
\end{CodeOutput}
\begin{CodeOutput}
AIC= 10833.841
\end{CodeOutput}
\begin{CodeOutput}
No survey area information supplied, only estimating detection function.
\end{CodeOutput}
\begin{CodeInput}
gof_ds(amakihi_hn)
\end{CodeInput}
\begin{CodeOutput}

Goodness of fit results for ddf object

Distance sampling Kolmogorov-Smirnov test
Test statistic =  0.059345  P =  0.00031527

Distance sampling Cramer-von Mises test (unweighted)
Test statistic =  0.93083  P =  0.003578 
\end{CodeOutput}
\begin{CodeInput}
gof_ds(amakihi_hr_obs_mas)
\end{CodeInput}
\begin{CodeOutput}

Goodness of fit results for ddf object

Distance sampling Kolmogorov-Smirnov test
Test statistic =  0.036251  P =  0.076237

Distance sampling Cramer-von Mises test (unweighted)
Test statistic =  0.15016  P =  0.38908 
\end{CodeOutput}
\end{CodeChunk}

We can conclude that the half-normal model does not pass our goodness of
fit tests and should be discarded. Both hazard-rate models (output only
shown for hazard-rate model with observer and minutes after sunrise) had
non-significant goodness of fit test statistics and are both therefore
deemed plausible models. The corresponding Q-Q plots are shown in Figure
\ref{amakihi-qq}, comparing the half-normal model with the hazard-rate
model with observer and minutes after sunrise included.

\begin{CodeChunk}
\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/amakihi-qq-comp-1}

}

\caption[Comparison of quantile-quantile plots for a half-normal model (no adjustments, no covariates]{Comparison of quantile-quantile plots for a half-normal model (no adjustments, no covariates; left) and hazard-rate model with observer and minutes after sunrise (right) for the amakihi data.\label{amakihi-qq}}\label{fig:amakihi-qq-comp}
\end{figure}
\end{CodeChunk}

For non-exact distance data, a \(\chi^2\)-test can be used to assess
goodness of fit \citep[see][Section 3.4.4]{Buckland:2001vm}.
\(\chi^2\)-test results are produced by \texttt{gof\_ds} when
binned/grouped distances are provided.

\subsection{Model selection}\label{model-selection}

Once we have a set of plausible models, we can use Akaike's Information
Criterion (AIC) to select between models \citep[see
e.g.][]{burnham2003model}. \pkg{Distance} includes a function to create
table of summary information for fitted models, making it easy to get an
overview of a large number of models. The \code{summarize_ds_models}
function takes models as input and can be especially useful when paired
with \pkg{knitr}'s \code{kable} function to create summary tables for
publication \citep{knitr-pkg}. An example of this output (with all
models included) is shown in Table \ref{tab:amakihi} and was generated
by the following call to \code{summarize_ds_models}:

\begin{CodeChunk}
\begin{CodeInput}
summarize_ds_models(amakihi_hn, amakihi_hr_obs, amakihi_hr_obs_mas)
\end{CodeInput}
\end{CodeChunk}

In this case we may be skeptical about the top model as selected by AIC
being truly better than the second best model, as there is only a very
small difference in AICs. Generally, if the difference between AICs is
less than 2, we may investigate multiple ``best'' models, potentially
resorting to the simplest of these models. In the authors' experience,
it is often the case that models with similar AICs will have similar
estimates probabilities of detection, so in practice there is little
difference in selecting between these models. It is important to note
that comparing AICs between models with different truncations is not
appropriate, as models with different truncation use different data.

\begin{table}

\centering
\begin{tabular}[t]{lllrrrr}
\toprule
Model & Key function & Formula & C-vM $p$-value & $\hat{P_a}$ & se($\hat{P_a}$) & $\Delta$AIC\\
\midrule
\texttt{amakihi\char`_hr\char`_obs\char`_mas} & Hazard-rate & ~obs + mas & 0.389 & 0.319 & 0.020 & 0.000\\
\texttt{amakihi\char`_hr\char`_obs} & Hazard-rate & ~obs & 0.271 & 0.314 & 0.020 & 1.073\\
\texttt{amakihi\char`_hn} & Half-normal & ~1 & 0.004 & 0.351 & 0.011 & 56.465\\
\bottomrule
\end{tabular}
\caption{Summary for the detection function models fitted to the amakihi data. ``C-vM'' stands for Cramér-von Mises, $P_a$ is average detectability (see ``Estimating abundance and variance''), se is standard error. Models are sorted according to AIC.\label{tab:amakihi}}
\end{table}

\section{Estimating abundance and
variance}\label{estimating-abundance-and-variance}

Though fitting the detection function is the primary modelling step in
distance sampling, we are really interested in estimating density or
abundance. It is also important to calculate our uncertainty for these
estimates. This section addresses these issues mathematically before
showing how to estimate abundance and its variance in \proglang{R}.

\subsection{Abundance}\label{abundance}

We wish to obtain the abundance in a study region, of which we have
sampled a random subset. To do this we first calculate the abundance in
the area we have surveyed (the \emph{covered area}) to obtain
\(\hat{N}_\text{C}\), we can then scale this up to the full study area
by multiplying it by the ratio of covered area to study area. We discuss
other methods for spatially explicit abundance estimation in
``Extensions''.

First, to estimate abundance in the covered area (\(\hat{N}_\text{C}\)),
we use the estimates of detection probability (the
\(\{\hat{p}(\mathbf{z}_i); i=1,\ldots,n\}\), above) in a
Horvitz-Thompson-like estimator:

\begin{equation}
\hat{N}_\text{C} = \sum_{i=1}^n\frac{s_i}{\hat{p}(\mathbf{z}_i)},
\label{ht}
\end{equation}

where \(s_i\) are the sizes of the observed clusters of objects, which
are all equal to 1 if objects only occur singly \citep{Borchers:2004wr}.
\citet{thompson2012sampling} is the canonical reference to this type of
estimator. Intuitively, we can think of the estimates of detectability
(\(\hat{p}(\mathbf{z}_i)\)) as ``inflating'' the cluster sizes
(\(s_i\)), we then sum over the detections (\(i\)) to obtain the
abundance estimate. For models that do not include covariates,
\(\hat{p}(\mathbf{z}_i)\) is equal for all \(i\), so this is equivalent
to summing the clusters and inflating that sum by dividing through by
the corresponding \(\hat{p} (=\hat{p}(\mathbf{z}_i) \forall i)\).

Having obtained the abundance in the covered area, we can then scale-up
to the study area: \[
\hat{N} = \frac{A}{a} \hat{N}_\text{C},
\] where \(A\) is the area of the study region to which to extrapolate
the abundance estimate and \(a\) is the covered area. For line transects
\(a=2wL\) (twice the truncation distance multiplied by the total length
of transects surveyed, \(L\)) and for points \(a=\pi w^2 T\) (where
\(\pi w^2\) is the area of a single surveyed circle and \(T\) is the sum
of the number of visits to the sampled points).

We can use the Horvitz-Thompson-like estimator to calculate the
``average'' detectability for models which include covariates. We can
consider what single detectability value would give the estimated
\(\hat{N}_\text{C}\) and therefore calculate: \[
\hat{P_a} = n/\hat{N}_\text{C}.
\] This can be a useful summary statistic, giving us an idea of how
detectable our \(n\) observed animals would have to be to estimate the
same \(\hat{N}\) if there were no observed covariates. It can also be
compared to similar estimates in mark-recapture studies. \(P_a\) is
included in the \code{summary} output and the table produced by
\code{summarize_ds_models}.

\subsubsection{Stratification}\label{stratification}

We may wish to calculate abundance estimates for some sub-regions of the
study region, we call these areas \emph{strata}. Stratification can be
used to increase the precision of estimates if we know \emph{a priori}
that density varies between different parts of the study area. For
example, strata may be defined by habitat types or animal sex (or some
combination) which may be of interest for biological or management
reasons. To calculate estimates for a given stratification each
observation must occur in a stratum which must be labelled with a
\code{Region.Label} and have a corresponding \code{Area} (if we are
using an animal characteristic like sex, we would have the areas be the
same but if we were using say forested vs.~wetland habitat the areas of
those strata would be different). Finally, we must also know the stratum
in which each observation occurs.

As an example, the minke whale data contain two strata: \code{North} and
\code{South} relating to strata further away from and nearer to the
Antarctic ice edge, respectively.

\subsection{Variance}\label{variance}

We take an intuitive approach to uncertainty estimation, for a full
derivation consult \citet{Marques:2003vb}. Uncertainty in \(\hat{N}\)
comes from two sources:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \emph{Detection function}: Uncertainty in parameter
  (\(\boldsymbol{\theta}\)) estimation.
\item
  \emph{Encounter rate}: Sampling variability arising from differences
  in the number of observations per transect.
\end{enumerate}

We can see this by looking at the Horvitz-Thompson-like estimation in
(\ref{ht}) and consider the terms which are random. These are: the
detectability \(\hat{p}(\mathbf{z}_i)\) (and hence the parameters of the
detection function it is derived from) and \(n\), the number of
observations.

Model parameter uncertainty can be addressed using standard maximum
likelihood theory. We can invert the Hessian matrix of the likelihood to
obtain a variance-covariance matrix. We can then pre- and post-multiply
this by the derivatives of \(\hat{N}_\text{C}\) with respect to the
parameters of the detection function \[
\widehat{\text{Var}}_\text{model}\left( \hat{N}_\text{C}\right) = \left(\frac{\partial \hat{N}_\text{C}}{\partial\hat{\boldsymbol{\theta}}}\right)^\text{T} \left(\hat{\mathbf{H}}(\hat{\boldsymbol{\theta}})^{-1} \right)\frac{\partial \hat{N}_\text{C}}{\partial\hat{\boldsymbol{\theta}}}
\] where the partial derivatives of \(\hat{N}_\text{C}\) are evaluated
at the MLE (\(\hat{\theta}\)) and \(\mathbf{H}\) is the first partial
Hessian (outer product of first derivatives of the log likelihood) for
numerical stability \citep[p 62]{Buckland:2001vm}. Note that although we
calculate uncertainty in \(\hat{N}_\text{C}\), we can trivially scale-up
to variance of \(\hat{N}\) (by noting that
\(\hat{N} = \frac{A}{a} \hat{N}_\text{C}\) and therefore
\(\widehat{\text{Var}}_\text{model}\left(\hat{N}\right) = \left( \frac{A}{a} \right)^2 \widehat{\text{Var}}_\text{model} \left(\hat{N}_\text{C} \right)\)).

\emph{Encounter rate} is the number of objects per unit transect (rather
than just \(n\)). When covariates are not included in the detection
function we can define the encounter rate as \(n/L\) for line transects
(where \(L\) is the total line length) or \(n/T\) for point transects
(where \(T\) is the total number of visits summed over all points). When
covariates are included in the detection function, it is recommended
that we substitute the \(n\) in the encounter rate with the estimated
abundance \(\hat{N}_\text{C}\) as this will take into account the
effects of the covariates \citep{Innes:2002ka}.

For line transects, by default, \pkg{Distance} uses a variation of the
estimator ``R2'' from \citet{Fewster:2009ku} replacing number of
observations per sample with the estimated abundance per sample, thus
taking into account cluster size if this is recorded
\citep{Innes:2002ka, Marques:2003vb}: \[
\widehat{\text{Var}}_{\text{encounter},R2}\left( \hat{N}_\text{C}\right) = \frac{K}{L^2(K-1)} \sum_{k=1}^{K} l_k^2 \left( \frac{\hat{N_{\text{C}, k}}}{l_k} - \frac{\hat{N_\text{C}}}{L}\right)^2,
\] where \(l_k\) are the lengths of the \(K\) transects (such that
\(L = \sum_{k=1}^K l_k\)) and \(\hat{N_{C,k}}\) is the abundance in the
covered area for transect \(k\). For point transects we use the
estimator ``P3'' from \citet{Fewster:2009ku} but again replace \(n\) by
\(\hat{N}_\text{C}\) in the encounter rate definition, to obtain the
following estimator: \[
\widehat{\text{Var}}_{\text{encounter},P3}\left( \hat{N}_\text{C}\right) = \frac{1}{T(K-1)} \sum_{k=1}^{K} t_k \left( \frac{\hat{N_{\text{C}, k}}}{t_k} - \frac{\hat{N_\text{C}}}{T}\right)^2,
\] where \(t_k\) is the number of visits to point \(k\) and
\(T = \sum_{k=1}^K t_k\) (the total number of visits to all points is
the sum of the visits to each point). Again, it is straightforward to
calculate the encounter rate variance for \(\hat{N}\) from the encounter
rate variance for \(\hat{N}_\text{C}\).

Other formulations for the encounter rate variance are discussed in
detail in \citet{Fewster:2009ku}. \pkg{Distance} implements all of the
estimators of encounter rate variance given in that article. The
\code{varn} manual page gives further advice and technical detail on
encounter rate variance. For example for systematic survey designs,
estimators S1, S2 and O1, O2 and O3 will typically provide smaller
estimates of the variance.

We combine these two variances by noting that squared coefficients of
variation (approximately) add \citep[often referred to as ``the delta
method'';][]{seber2002estimation}.

\subsection{Estimating abundance and variance in
R}\label{estimating-abundance-and-variance-in-r}

Returning to the minke whale data, we have the necessary information to
calculate \(A\) and \(a\) above, so we can estimate abundance and its
variance. When we supply data to \code{ds} in the ``flatfile'' format
given above, \code{ds} will automatically calculate abundance estimates
based on the survey information in the data.

Having already fitted a model to the minke whale data, we can see the
results of the abundance estimation by viewing the model summary:

\begin{CodeChunk}
\begin{CodeInput}
summary(minke_hn)
\end{CodeInput}
\begin{CodeOutput}

Summary for distance analysis 
Number of observations :  88 
Distance range         :  0  -  1.5

Model : Half-normal key function 
AIC   : 46.87216

Detection function parameters
Scale coefficient(s):  
              estimate        se
(Intercept) -0.3411766 0.1070304

                       Estimate          SE         CV
Average p             0.5733038  0.04980421 0.08687229
N in covered region 153.4962706 17.08959835 0.11133559

Summary statistics:
  Region   Area CoveredArea  Effort  n  k         ER      se.ER     cv.ER
1  North 630582     4075.14 1358.38 49 12 0.03607238 0.01317937 0.3653591
2  South  84734     1453.23  484.41 39 13 0.08051031 0.01809954 0.2248102
3  Total 715316     5528.37 1842.79 88 25 0.04775368 0.01129627 0.2365529

Abundance:
  Label Estimate        se        cv      lcl      ucl       df
1 North 13225.44 4966.7495 0.3755450 6005.590 29124.93 12.27398
2 South  3966.46  955.9616 0.2410113 2395.606  6567.36 15.80275
3 Total 17191.90 5135.5862 0.2987212 9183.475 32184.07 14.00459

Density:
  Label   Estimate          se        cv         lcl        ucl       df
1 North 0.02097339 0.007876453 0.3755450 0.009523884 0.04618738 12.27398
2 South 0.04681073 0.011281913 0.2410113 0.028272077 0.07750560 15.80275
3 Total 0.02403400 0.007179465 0.2987212 0.012838347 0.04499280 14.00459
\end{CodeOutput}
\end{CodeChunk}

This prints a rather large amount of information: first the detection
function summary, then three tables:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \code{Summary statistics}: giving the areas, covered areas, effort,
  number of observations, number of transects, encounter rate, its
  standard error and coefficient of variation for each stratum.
\item
  \code{Abundance}: giving estimates, standard errors, coefficients of
  variation, lower and upper confidence intervals and finally the
  degrees of freedom for each stratum's abundance estimate.
\item
  \code{Density}: lists the same statistics as \code{Abundance} but for
  a density estimate.
\end{enumerate}

In each table the bottom row gives a total for the whole study area.

The summary can be more concisely expressed by extracting information
from the summary object. This object is a \code{list} of
\code{data.frames}, so we can use the \code{kable} function from
\pkg{knitr} to create summary tables of abundance estimates and measures
of precision, such as Table \ref{minke-abund}. We prepare the
\code{data.frame} as follows before using \code{kable}:

\begin{CodeChunk}
\begin{CodeInput}
minke_table <- summary(minke_hn)$dht$individuals$N
minke_table$lcl <- minke_table$ucl <- minke_table$df <- NULL
colnames(minke_table) <- c("Stratum", "$\\hat{N}$", "$\\text{se}(\\hat{N}$)", 
                           "$\\text{CV}(\\hat{N}$)")
\end{CodeInput}
\end{CodeChunk}

\begin{CodeChunk}
\begin{table}

\centering
\begin{tabular}[t]{lrrr}
\toprule
Stratum & $\hat{N}$ & $\text{se}(\hat{N}$) & $\text{CV}(\hat{N}$)\\
\midrule
North & 13225.44 & 4966.750 & 0.376\\
South & 3966.46 & 955.962 & 0.241\\
Total & 17191.90 & 5135.586 & 0.299\\
\bottomrule
\end{tabular}
\caption{Summary of abundance estimation for the half-normal model for the minke whale data.\label{minke-abund}}
\end{table}

\end{CodeChunk}

\section{Extensions}\label{extensions}

Distance sampling has been applied in a wide variety of situations.
Objects being detected need not be animals; plants
\citep{Buckland:2007kl}, beer cans \citep{Otto:1990gs}, elephant dung
\citep{NchPlu01} and bricks on a lake bottom \citep{Ber90} have been
subjects of distance sampling investigations. The method of detection
need not be visual sightings. Detections of objects can be through
auditory as well as visual means \citep{Marques2013b}. Songs of birds
\citep{Buckland2006} and whale vocalisations \citep{BorSub1} are just
two examples. Blows made by whales are examples of processes that can be
modelled using distance sampling. Songs and blows are indirect sampling
methods producing estimates of cue density. Cue densities can be
converted to animal densities with additional data needed to estimate
the rate at which cues are produced and the rate at which they
disappear. These are but a few examples of the applications to which
distance sampling has been applied (a larger, but incomplete list of
references is given at \url{http://distancesampling.org/dbib.html}).

The features of \pkg{Distance} are deliberately limited to provide a
simplified interface for users. For more complex analyses of distance
sampling data, we provide additional packages for modelling in
\proglang{R}.

We noted at the start of the article that \pkg{Distance} is a
simple-to-use wrapper around the package \pkg{mrds}. Additional features
available in \pkg{mrds} include models that relax the assumption that
detection is certain at zero distance from the transect (by including
data from additional observers). This is done using mark-recapture type
methods which require additional survey methodology, known as double
observer surveys \citep[see][ for an introduction]{Burt:2014gu}.

\pkg{Distance} can provide us with estimates of abundance or density for
each strata as a whole but tells us nothing about the distribution of
animals within strata. One option is to divide the study area into
smaller and smaller strata to try to detect patterns in spatial
distribution, however, a more rigorous approach is to build a spatial
model. Such models incorporate spatially-referenced environmental data
(for example derived from GIS products). \pkg{Distance} interfaces with
one such package used to perform this type of analysis: \pkg{dsm}
\citep{dsm-pkg}. So-called ``density surface modelling'' uses the
generalized additive model framework \citep[e.g.][]{Wood:2006vg} to
build models of abundance (adjusting counts for imperfect detectability)
as a function of environmental covariates, as part of a two stage model
\citep{Hedley:2004et, Miller:2013fq}.

Uncertainty in measured covariates (e.g.~cluster size) and model
uncertainty (when two models have similar fit but substantially
different estimates) can be incorporated using the multi-analysis
distance sampling package \pkg{mads} \citep{mads-pkg}. In addition,
\pkg{mads} can also incorporate sightings with unknown species
identification. This is done by estimating the abundance of these
`unidentified' sightings and pro-rating them to the known species
\citep{Gerrodette:2005}.

As mentioned above, survey design is critical to ensuring that resulting
distance sampling data can be reliably analysed. \pkg{DSsim} allows
users to test out different designs in their study region and tailor
population attributes to reflect the species they are working with.
\pkg{DSsim} \citep{DSsim-pkg} allows users to more easily identify
challenges unique to their study and select a survey design which is
more likely to yield accurate and precise estimates.

Distance for Windows has many users (over 45,000 downloads since 2002)
and they may be overwhelmed by the prospect of switching existing
analyses to \proglang{R}. For that reason we have created the
\pkg{readdst} \citep{readdst-pkg} package to interface with projects
created by Distance for Windows. The package can take analyses created
using the CDS, MCDS and MRDS engines in Distance for Windows, extract
the data and create equivalent models in \proglang{R}. \pkg{readdst} can
also run these analyses and test the resulting statistics (for example,
\(\hat{N}\) or \(\hat{P_a}\)) calculated in \proglang{R} against those
calculated by Distance for Windows. We hope that \pkg{readdst} will
provide a useful transition to \proglang{R} for interested users.
\pkg{readdst} is currently available on GitHub at
\url{https://github.com/distancedevelopment/readdst}.

\section{Conclusion}\label{conclusion}

We have given an introduction as to how to perform a distance sampling
analysis in \proglang{R}. We have covered the possible models for
detectability, model checking and selection and finally abundance and
variance estimation.

In combination with tools such as \pkg{knitr} and \pkg{rmarkdown}
\citep{rmarkdown-pkg}, the helper functions in \pkg{Distance} provide a
useful set of tools to perform reproducible analyses of wildlife
abundance for both managers and ecologists. We also direct readers'
attention to \pkg{DsShiny} \citep{dsshiny-pkg}, a package that builds on
\pkg{Shiny} \citep{shiny-pkg} and \pkg{mrds} to allow users to build and
fit models in a graphical interface, with output to RMarkdown.

\proglang{R} and its extension packages provide many tools exploratory
data analysis that can be useful for a distance sampling analysis. We
hope that this paper provides useful examples for those wishing to
pursue distance sampling in \proglang{R}. More information on distance
sampling can be found at \url{http://distancesampling.org} and a mailing
list is maintained at
\url{https://groups.google.com/forum/#!forum/distance-sampling}.

We note that there are other packages available for performing distance
sampling analyses in \proglang{R} but believe that \pkg{Distance} is the
most flexible and feature-complete, and provides pathways to a range of
more complicated analyses. Appendix A gives a feature comparison between
\pkg{Distance} and other \proglang{R} packages for analysis of distance
sampling data.

\section{Acknowledgements}\label{acknowledgements}

The authors would like to thank the many users of \pkg{Distance},
\pkg{mrds} and Distance for Windows who have contributed bug reports and
suggestions for improvements over the years. We would particularly like
to thank Steve Buckland, David Borchers, Tiago Marques, Jon Bishop and
Lorenzo Milazzo for their contributions. We thank Tiago Marques a second
time for helpful comments on an early version of the paper and Colin
Beale who suggested the ``flatfile'' data format. We also thank Ken
Burnham and David Anderson for fundamental contributions to the early
development of these methods.

\renewcommand\refname{Bibliography}
\bibliography{jstatsoft.bib}

\end{document}