### 2016-04-05 appendix-A.Rmd
appendix A

---
author:
  - name: David L Miller
    affiliation: University of St Andrews
  - name: Eric Rexstad
    affiliation: University of St Andrews
  - name: Len Thomas
    affiliation: University of St Andrews
  - name: Laura Marshall
    affiliation: University of St Andrews
  - name: Jeffrey L Laake
    affiliation: Marine Mammal Laboratory
title: "Appendix A: Feature comparison between \\pkg{Distance} and other packages"
preamble: >
  \usepackage{amsmath}
  \usepackage{amssymb}
  \usepackage{longtable}
  \usepackage{booktabs}
output: html_document
bibliography: jstatsoft.bib

---

There are four packages available for analysis of distance sampling data in R that we are aware of. All are available on CRAN.

Package  | Description  | Version  | Homepage
-------: | -----------: | -------: | ---------:
`Distance` | package described in the main paper | 0.9.6 | https://github.com/distancedevelopment/Distance
`Rdistance` | uses a similar approach to `Distance` | 1.3.2 | https://github.com/tmcd82070/Rdistance
`unmarked` | uses a heirarchical modelling approach | 0.11.0 | https://sites.google.com/site/unmarkedinfo/
`mrds`     | on which `Distance` is based  | 2.1.15 | https://github.com/distancedevelopment/mrds

This table provides a feature comparison of the packages:

Feature    |  `Distance` | `Rdistance` | `unmarked` | `mrds`
-------: | --------:  | ---------: | --------: | ----:
Line transects | x | x | x | x
Point transets | x | | x | x
Interval (binned) distances | x | | x | x
Exact distances | x | x | | x
Continuous individual level covariates | x |  | | x
Factor individual level covariates | x | x | | x
Transect level covariates | x |  | x | x
Objects in clusters | x | x | x | x
Left truncation | x  | x | | x
Half-normal key | x | x | x | x
Hazard-rate key | x | x |  x | x
Uniform key | x | x | x | x
Gamma key |  | x |  | x
Negative exponential key |  |  | x  |
Adjustment terms | x | x | | x
AIC adjustment selection | x | x | 
Monotonicity constraints | x |  | | x
Availability bias model | | | x |
Perception bias model | | | | x
Abundance estimation | x | x | x | x
Density estimation | x | x | x | x
User-defined likelihood functions | | x | |

## Abundance estimation

`Distance` and `mrds` models can be used as part of a density surface model using `dsm`, which allows abundance to be modelled as a function of spatially varying covariates (such as location, sea depth, altitude etc). See @Miller:2013fq for more information.

`unmarked` also allows abundance to vary accoring to covariates, via the abundance part of the likelihood. See @Fiske:2011gv for more information on the package and @Royle:2004cd for more information on methodology.

Results `Rdistance` models can be use in combination with R modelling functions such as `lm`, `glm` etc to build abundance estimates which vary according to covariates. More information is available on the project's wiki.

# References